# Enrichment Analysis 

`r if(enrichr_section&length(gene_data_index)>0){"### EnrichR"}`
```{r enrichr, eval=(enrichr_section&length(gene_data_index)>0) }

output_links<-"";
run_enrichr_seperate <- TRUE;

enrichr_working_dir <- "EnrichR"
enrichr_working_path <- file.path(output_path, enrichr_working_dir)
if( dir.exists(enrichr_working_path) == FALSE ) { dir.create(enrichr_working_path) }
  
# Run Enrichr for each data type
for(i in 1:length(gene_data_index) ){
  sig_cutoff <- round(sig_percent*nrow(omicsList[[ gene_data_index[i] ]][["eSet"]]), digits=0)
  for(c in 1:length(loop_list) ){ try({
    top_sum<- FALSE;
    contrast_name <- contrast_strings[c]
    try({
      top_sum <- topTable(omicsList[[ gene_data_index[i] ]][["fit"]], adjust="BH", n=Inf, sort.by='p', coef=loop_list[c], p.value=adjpcutoff);
      if (nrow(top_sum)<sig_cutoff){ top_sum <- topTable(omicsList[[ gene_data_index[i] ]][["fit"]], adjust="BH", n=sig_cutoff, sort.by='p', coef=loop_list[c]); }
    })
    if( class(top_sum)=="logical" ){
      top_sum <- data.frame(logFC= fData(omicsList[[ gene_data_index[i] ]][["eSet"]])[,paste("logfc_",gsub("-","_",contrast_strings[c]),sep="")],
                            Gene=fData(omicsList[[ gene_data_index[i] ]][["eSet"]])$Gene, 
                            feature_identifier =fData(omicsList[[ gene_data_index[i] ]][["eSet"]])[,"feature_identifier"] );
      top_sum <- top_sum[top_sum[,"Gene"]!="", ]
      top_sum <- top_sum[order(abs(as.numeric(top_sum[,"logFC"])), decreasing=TRUE),]
      if( nrow(top_sum)>sig_cutoff ){  top_sum <- top_sum[1:sigcutoff,] }
      contrast_name <- gsub("logfc_","",logfc_index[c])
    }
    type_name <- paste(omicsList[[ gene_data_index[i] ]][["dataType"]], contrast_name, sep="_");
    
    runEnrichR(genes=top_sum[, c("Gene", "logFC") ], type=type_name, run_seperate=run_enrichr_seperate) 
    
    # Make output hyperlinks for markdown
    add_link <- paste("[ ",type_name, " ](", enrichr_working_dir,"/enrichr_",type_name,".txt) | ", sep="")
    output_links <- paste(output_links, add_link, sep="  \n" )
  }) } 
  
  # Timecourse
  if(time_index>0){try({
    top_sum<- FALSE;
    contrast_name <- "TimeCourse_Overall"
    top_sum <- topTable(omicsList[[ gene_data_index[i] ]][["fit"]], adjust="BH", n=Inf, coef=time_start:time_end, p.value=adjpcutoff);
    if (nrow(top_sum)<sig_cutoff){ top_sum <- topTable(omicsList[[ gene_data_index[i] ]][["fit"]], adjust="BH", n=sig_cutoff, coef=time_start:time_end); }
    top_sum[,"logFC"] <- top_sum[,'F']
    type_name <- paste(omicsList[[ gene_data_index[i] ]][["dataType"]], "TimeCourse_Overall", sep="_");
    
    runEnrichR(genes=top_sum[, c("Gene", "logFC") ], type=type_name, run_seperate=run_enrichr_seperate) 
    
    # Make output hyperlinks for markdown
    add_link <- paste("[ ",type_name, " ](", enrichr_working_dir,"/enrichr_",type_name,".txt) | ", sep="")
    output_links <- paste(output_links, add_link, sep="  \n" )
  }) } 
} 
  
# Run enrichr for combined data
try({
if( length(gene_data_index)>1 ){
  sig_cutoff <- round(sig_percent*nrow(omicsList[[ gene_data_index[1] ]][["eSet"]]), digits=0)
  for(c in 1:length(loop_list)){ 
    type_name <- paste("Combined",contrast_strings[c], sep="_");
    gene_table <- topTable(omicsList[[ gene_data_index[1] ]][["fit"]], adjust="BH", n=Inf, sort.by='p', coef=loop_list[c], p.value=adjpcutoff)
    if (nrow(gene_table)<sig_cutoff){ gene_table<-topTable(omicsList[[ gene_data_index[1] ]][["fit"]], adjust="BH",
                                                           n=sig_cutoff, sort.by='p', coef=loop_list[c]) }
    gene_table <- gene_table[, c("Gene", "logFC") ];
        
    for(i in 2:length(gene_data_index) ){
      sig_cutoff <- round(0.05*nrow(omicsList[[ gene_data_index[i] ]][["eSet"]]), digits=0)
      add_to_table <- topTable(omicsList[[ gene_data_index[i] ]][["fit"]], adjust="BH", n=Inf,
                               sort.by='p', coef=loop_list[c], p.value=adjpcutoff);
      if (nrow(add_to_table)<sig_cutoff){ add_to_table<-topTable(omicsList[[ gene_data_index[i] ]][["fit"]], adjust="BH", n=sig_cutoff, sort.by='p', coef=loop_list[c]); }
      add_to_table <- add_to_table[, c("Gene", "logFC") ];
      gene_table <- rbind(gene_table, add_to_table)
    }
    
    runEnrichR(gene_table[, c("Gene", "logFC") ], type=type_name, run_seperate=run_enrichr_seperate )
   
    # Make output hyperlinks for markdown
    add_link <- paste("[ ",type_name, " ](", enrichr_working_dir,"/enrichr_",type_name,".txt) | ", sep="")
    output_links <- paste(output_links, add_link, sep="  \n" )
  } 
   if(statistic_index=='F'){ 
    sig_cutoff <- round(sig_percent*nrow(omicsList[[ gene_data_index[1] ]][["eSet"]]), digits=0)
    type_name <- paste("Combined","Overall", sep="_");
    gene_table <- topTableF(omicsList[[ gene_data_index[1] ]][["fit"]], adjust="BH", n=Inf, sort.by='F', p.value=adjpcutoff)
    if (nrow(gene_table)<sig_cutoff){ gene_table<-topTableF(omicsList[[ gene_data_index[1] ]][["fit"]], adjust="BH", n=sig_cutoff, sort.by='F') }
    gene_table <- gene_table[, c("Gene", "F") ];
        
    for(i in 2:length(gene_data_index) ){
      sig_cutoff <- round(sig_percent*nrow(omicsList[[ gene_data_index[i] ]][["eSet"]]), digits=0)
      add_to_table <- topTableF(omicsList[[ gene_data_index[i] ]][["fit"]], adjust="BH", n=Inf, sort.by='F', p.value=adjpcutoff);
      if (nrow(add_to_table)<sig_cutoff){ add_to_table<-topTableF(omicsList[[ gene_data_index[i] ]][["fit"]], adjust="BH", n=sig_cutoff, sort.by='F'); }
      add_to_table <- add_to_table[, c("Gene", "F") ];
      gene_table <- rbind(gene_table, add_to_table)
    }
    gene_table[,"logFC" ] <- gene_table[,"F" ] 
    runEnrichR(gene_table[, c("Gene", "logFC") ], type=type_name, run_seperate=run_enrichr_seperate )
   
    # Make output hyperlinks for markdown
    add_link <- paste("[ ",type_name, " ](", enrichr_working_dir,"/enrichr_",type_name,".txt) | ", sep="")
    output_links <- paste(output_links, add_link, sep="  \n" )
  } 
} 
})

# Add enrichr based on F stat
for(i in 1:length(gene_data_index) ){ 
  if(statistic_index =='F'){ try({
    top_sum<- FALSE;
    contrast_name <- "F-statistic"
    sig_cutoff <- round(sig_percent*nrow(omicsList[[ gene_data_index[1] ]][["eSet"]]), digits=0)
    top_sum <- topTableF(omicsList[[ gene_data_index[i] ]][["fit"]], adjust="BH", n=Inf, sort.by='F', p.value=adjpcutoff);
    if (nrow(top_sum)<sig_cutoff){ top_sum <- topTableF(omicsList[[ gene_data_index[i] ]][["fit"]], adjust="BH", n=sig_cutoff, sort.by='F'); }
    top_sum[,"logFC"] <- top_sum[,'F']
    type_name <- paste(omicsList[[ gene_data_index[i] ]][["dataType"]], 'F-statistic', sep="_");
    
    runEnrichR(genes=top_sum[, c("Gene", "logFC") ], type=type_name, run_seperate=run_enrichr_seperate) 
    
    # Make output hyperlinks for markdown
    add_link <- paste("[ ",type_name, " ](", enrichr_working_dir,"/enrichr_",type_name,".txt) | ", sep="")
    output_links <- paste(output_links, add_link, sep="  \n" )
  }) }
} 

```
`r if(enrichr_section&length(gene_data_index)>0 & output_links!=""){"Click below to view enrichR results:  "}`  
`r if(enrichr_section&length(gene_data_index)>0){ output_links } `   

---
```{r echo=FALSE}
enrichment_map <- gsea_section; # only run enrichment map if interactive session
output_network_file <- FALSE;
```

`r if( gsea_section & species!="Other" & length(gene_data_index)>0){"### GSEA and EnrichmentMap"} `
```{r gsea_section, eval=(gsea_section & length(gene_data_index)>0) }
## This sections is based on the Bader Lab Enrichment Map Workflow.
## The connection between R and cytoscape must be configured for proper functioning.
## For more info: baderlab.github.io/Cytoscape_workflows/EnrichmentMapPipeline/

output_links <- "";
run_gsea <- gsea_section; # if GSEA has already been run, set to false and include results in directory
enrichment_map <- gsea_section; # only run enrichment map if interactive session

if( run_gsea & species!="Other" ) { try({
  
  gsea_working_dir <- "GSEA"
  gsea_working_path <- file.path(output_path, gsea_working_dir)
  gsea_images_path <- file.path(gsea_working_path, "Images")
  gsea_absval_path <- file.path(gsea_working_path, "AbsoluteValueRank")
  if( dir.exists(gsea_working_path) == FALSE ) { dir.create(gsea_working_path) }
  if( dir.exists(gsea_images_path) == FALSE ) { dir.create(gsea_images_path) }
  if( dir.exists(gsea_absval_path) == FALSE ) { dir.create(gsea_absval_path) }
  
    analysis_names <- vector("list", (length(loop_list)*length(gene_data_index)) ) ;
    counter<-1;                       
    for (j in 1:length(gene_data_index)) {
      for (k in 1:length(loop_list)) {
        analysis_names[[counter]] <- paste(omicsList[[ gene_data_index[j] ]][["dataType"]], "_", gsub("-", "_", contrast_strings[k]), sep="") ;
        counter <- counter+1;
      }
    }
    rnk_files <- vector("list", (length(loop_list)*length(gene_data_index)) ) ;
    counter<-1;                       
    for (j in 1:length(gene_data_index)) {
      for (k in 1:length(loop_list)) {
        if(grepl("Mouse", species) | grepl("Human", species)){ 
          rnk_files[[counter]] <- file.path(paste(output_contrast_path_files,"/GSEA_",omicsList[[ gene_data_index[j] ]][["dataType"]],
                                                  "_",gsub("-","_",contrast_strings[k]),".rnk", sep=""))
        } else {
          rnk_files[[counter]] <- file.path(paste(output_contrast_path_files,"/GSEA_",
                                                  omicsList[[ gene_data_index[j] ]][["dataType"]],"_Uppercase_",
                                                  gsub("-","_",contrast_strings[k]),".rnk", sep=""))
        }
        counter <- counter+1;
      }
    }
    expression_files <- vector("list", (length(loop_list)*length(gene_data_index)) ) ;
    counter<-1;                       
    for (j in 1:length(gene_data_index)) {
      for (k in 1:length(loop_list)) {
        expression_files[[counter]] <- file.path(paste(output_files_subdir, "/Expression_matrix_",
                                                       omicsList[[ gene_data_index[j] ]][["dataType"]],".txt", sep=""))
        counter <- counter+1;
      }
    } 
   if(length(gene_data_index)>1 & length(contrastgroups)>1){ 
    for (i in 1:length(loop_list)) {
      analysis_names[[length(analysis_names)+1]] <- paste("Combined_",gsub("-", "_", contrast_strings[i]),sep="")
    }
    for (i in 1:length(loop_list)){
        if(grepl("Mouse", species) | grepl("Human", species)){ 
          rnk_files <- append(rnk_files, file.path(output_contrast_path_files,
                                                        paste("GSEA_combined_",gsub("-","_",contrast_strings[i]),".rnk", sep="") ))
        } else {
          rnk_files <- append(rnk_files, file.path(output_contrast_path_files,
                                                        paste("GSEA_combined_Uppercase_",gsub("-","_",contrast_strings[i]),".rnk", sep="") ))
        }
    }
    for (i in 1:length(loop_list)){  
        expression_files <- append(expression_files, file.path(output_files_subdir,paste("Expression_matrix_",
                                                                                              omicsList[[ gene_data_index[1] ]][["dataType"]],
                                                                                              ".txt", sep="")))
    }
  }
  
  if(grepl("Mouse", species) | grepl("Human", species)){  suppressWarnings({ suppressMessages({
    # Only if you need a new GMT file
    if(grepl("Mouse", species)){gmt_url = "http://download.baderlab.org/EM_Genesets/current_release/Mouse/symbol/"}
    if(grepl("Human", species)){gmt_url = "http://download.baderlab.org/EM_Genesets/current_release/Human/symbol/"}
    filenames = getURL(gmt_url)   #list all the files on the server
    tc = textConnection(filenames)
    contents = readLines(tc)
    close(tc)
    #get the gmt that has all the pathways and does not include terms inferred from electronic annotations(IEA), start with gmt file that has pathways only
    rx = gregexpr("(?<=<a href=\")(.*.GOBP_AllPathways_no_GO_iea.*.)(.gmt)(?=\">)",contents, perl = TRUE)
    gmt_file = unlist(regmatches(contents, rx))
    dest_gmt_file <- file.path(gsea_working_path,gmt_file)
    if(!file.exists(dest_gmt_file)) { download.file(paste(gmt_url,gmt_file,sep=""),destfile=dest_gmt_file) }
  }) }) } else { try({
    geneset_lookup <-  read.delim(file.path(gsub("src", "data", notebook_dir),"geneset_table.txt"))
    gmt_file <- as.character(geneset_lookup[geneset_lookup[,"Species"]==species,"GeneSet"])
    dest_gmt_file <-file.path(gsea_working_path, gmt_file)
    file.copy(from=file.path( gsub("src", "data", notebook_dir), "species_genesets",gmt_file), to= dest_gmt_file )  
  }) }
  
  for (i in 1:length(analysis_names)){
    analysis_name <- analysis_names[[i]]
    rnk_file <- rnk_files[[i]]
    expression_file <- expression_files[[i]]
    
    # command <- paste("java  -Xmx1G -cp",gsea_jar,  "xtools.gsea.GseaPreranked -gmx",
    #                  dest_gmt_file, "-rnk" , rnk_file,
    #                  #"-cls",paste(working.dir,cls_filename,sep="/"),
    #                  "-collapse false -nperm 1000 -permute gene_set -scoring_scheme weighted -rpt_label",
    #                  analysis_name,"  -num 100 -plot_top_x 20 -rnd_seed 12345  -set_max 500 -set_min 15", 
    #                  "-zip_report false -out" ,gsea_working_path, "-gui false > ",
    #                  paste(gsea_working_path,"gsea_output.txt", sep="/"),sep=" ")
    # system(command)
    # # Clean up default directory
    # tmp<-list.files(working_dir) # get list of files
    # tmp<-tmp[file.info(tmp)$isdir==TRUE] # filter to just directories
    # tmp<-tmp[nchar(tmp)==5] # check pattern
    # tmp<-tmp[!is.na(tmp)] # remove na
    # if(length(tmp)==1){
    #   if(nchar(tmp[1])==5){file.remove(paste(working_dir, tmp[1], sep="/")) }
    # } 
    
    ranked_features <- read.delim(rnk_file)
    file.copy(from=rnk_file, to=file.path(gsea_working_path,basename(rnk_file)) )
    ranked_vector <- ranked_features[,"rank"]
    names(ranked_vector) <- ranked_features[,"GeneName"]
    pathways_gmt <- gmtPathways(dest_gmt_file)
    suppressWarnings({
    fgsea_results <- fgsea(pathways=pathways_gmt, 
                                     stats=ranked_vector,
                                     minSize=15, 
                                     maxSize=500, 
                                     nperm=10000)
    })
    
    fgsea_out <- fgsea_results[,c(1,1,2,3)]
    colnames(fgsea_out) <- c("Term", "Description", "p.Val", "FDR")
    fgsea_out[,"Phenotype"] <- sign(fgsea_results[,"NES"])
    fgsea_out[,"Genes"] <- apply(fgsea_results[,"leadingEdge"], 1, function(x) paste(as.character(unlist(x["leadingEdge"])), collapse=","))
    fgsea_out[,"Genes"] <- apply(fgsea_out, 1, function(x) gsub(" ", "", x["Genes"]))
    fgsea_out[,"NES"] <- (fgsea_results[,"NES"])
    fgsea_out[,"ES"] <- (fgsea_results[,"ES"])
    fgsea_out[,"Gene_Hits"] <- apply(fgsea_results, 1, function(x) length(unlist(x["leadingEdge"])) )
    fgsea_out[,"Gene_Total"] <- fgsea_results[,"size"]
    output_filename <- file.path(output_path, gsea_working_dir, paste("fgsea_", analysis_name, ".txt", sep=""))
    write.table(fgsea_out, file=output_filename, sep="\t", row.names=FALSE, quote=FALSE);
    
    output_filename <- file.path(gsea_images_path, paste("fgsea_", analysis_name, ".pdf", sep=""))
    topPathwaysUp <- fgsea_results[ES > 0][head(order(pval), n=10), pathway]
    topPathwaysDown <- fgsea_results[ES < 0][head(order(pval), n=10), pathway]
    topPathways <- c(topPathwaysUp, rev(topPathwaysDown))
    pdf(output_filename, width=12, height=8)
    plotGseaTable(pathways_gmt[topPathways], ranked_vector, fgsea_results, gseaParam = 0.5)
    dev.off()
    
    topPathways <- fgsea_results[order(pval, decreasing=F), pathway]
    if( length(topPathways) > 100 ) { topPathways <- topPathways[1:100]}
    output_filename <- file.path(gsea_images_path, paste("fgsea_", analysis_name, "_allPathways.pdf", sep=""))
    pdf(output_filename, width=4, height=3)
    for(pathway_index in 1:length(topPathways)){ try({
      print(plotEnrichment(pathways_gmt[[topPathways[pathway_index] ]], ranked_vector) +
              labs(title=topPathways[pathway_index]) + theme(plot.title=element_text(size=6)) )
    }, silent=TRUE) }
    dev.off()
    
    add_link <- paste("[ fgsea-",analysis_name, " ](", output_subdir,"/",gsea_working_dir,"/Images/fgsea_", analysis_name,".pdf)", sep="");
    output_links <- paste(output_links, add_link, sep=" | " ); 
    
    if(runDifferential==TRUE){
    ranked_features <- read.delim(rnk_file)
    ranked_vector <- abs(ranked_features[,"rank"])
    names(ranked_vector) <- ranked_features[,"GeneName"]
    pathways_gmt <- gmtPathways(dest_gmt_file)
    suppressWarnings({
    fgsea_results <- fgsea(pathways=pathways_gmt, 
                                     stats=ranked_vector,
                                     minSize=15, 
                                     maxSize=500, 
                                     nperm=10000)
    })
    fgsea_out <- fgsea_results[,c(1,1,2,3)]
    colnames(fgsea_out) <- c("Term", "Description", "p.Val", "FDR")
    fgsea_out[,"Phenotype"] <- sign(fgsea_results[,"NES"])
    fgsea_out[,"Genes"] <- apply(fgsea_results[,"leadingEdge"], 1, function(x) paste(as.character(unlist(x["leadingEdge"])), collapse=","))
    fgsea_out[,"Genes"] <- apply(fgsea_out, 1, function(x) gsub(" ", "", x["Genes"]))
    fgsea_out[,"NES"] <- (fgsea_results[,"NES"])
    fgsea_out[,"ES"] <- (fgsea_results[,"ES"])
    fgsea_out[,"Gene_Hits"] <- apply(fgsea_results, 1, function(x) length(unlist(x["leadingEdge"])) )
    fgsea_out[,"Gene_Total"] <- fgsea_results[,"size"]
    output_filename <- file.path(gsea_absval_path, paste("fgsea_", analysis_name, "_AbsVal.txt", sep=""))
    write.table(fgsea_out, file=output_filename, sep="\t", row.names=FALSE, quote=FALSE);
    }
  }

}) } 

# if(enrichment_map & species!="Other" ){
#   gsea_output_paths <- rev(row.names(details)[1:length(analysis_names)])
# 
#   try({
#   # Basic settings
#   port.number = 1234
#   base.url = paste("http://localhost:", toString(port.number), "/v1", sep="")
#   #print(base.url)
#   version.url = paste(base.url, "version", sep="/")
#   cytoscape.open = TRUE
#   tryCatch(expr = { GET(version.url)}, 
#          error = function(e) { return (cytoscape.open = FALSE)},
#          finally =function(r){ return(cytoscape.open = TRUE)})
#   
#   if(!cytoscape.open){
#   #try and launch cytoscape
#     print("Cytoscape is not open.  Please launch cytoscape.")
#   } else{
#     cytoscape.version =  GET(version.url)
#     cy.version = fromJSON(rawToChar(cytoscape.version$content))
#     print(cy.version)
#   } 
# 
#   #defined threshold for GSEA enrichments (need to be strings for cyrest call)
#   pvalue_gsea_threshold <- enrichmentmap_p_val
#   qvalue_gsea_threshold <- enrichmentmap_q_val
#   similarity_threshold <- "0.375"
#   similarity_metric = "COMBINED"
# 
#   cur_model_name <- analysis_names[1]
#   current_network_name <- paste(cur_model_name,pvalue_gsea_threshold,qvalue_gsea_threshold,sep="_")
#   
#   gsea_results_paths <- paste(gsea_output_paths,"edb",sep="/")
#   gsea_results_filenames <- paste(gsea_results_paths,"results.edb",sep="/")
# 
#   # Although there is a gmt file in the gsea edb results directory it have been filtered to contain only genes represented in the expression set.  
#   # If you use this fltered file you will get different pathway connectivity depending on the dataset being used.  
#   # We recommend using original gmt file used for the gsea analysis and not the filtered one in the results directory.
#   gmt_gsea_file <- paste(dest_gmt_file,sep="/")
# 
#   em_command = paste('enrichmentmap build analysisType="gsea" gmtFile=',gmt_gsea_file,
#                    'pvalue=',pvalue_gsea_threshold, 'qvalue=',qvalue_gsea_threshold,
#                    'similaritycutoff=',similarity_threshold,
#                    'coefficients=',similarity_metric,
#                    'ranksDataset1=',rnk_files[[1]],
#                    'enrichmentsDataset1=',gsea_results_filenames[1], 'filterByExpressions=false',
#                    'expressionDataset1=',file.path(working_dir,expression_files[[1]]),
#                    sep=" ")
# 
#   #enrichment map command will return the suid of newly created network.
#   try({ response <- commandsGET(em_command) }, silent=TRUE);
# 
#   current_network_suid <- 0
#   #enrichment map command will return the suid of newly created network unless it Failed.  If it failed it will contain the word failed
#   tryCatch({
#     if(length(response)>0){ #fix for not getting suid
#       if(grepl(pattern="Failed", response)){
#         paste(response)
#       } else {
#         current_network_suid <- response
#         response <- renameNetwork(new.name = current_network_name, network = current_network_suid, base.url)
#   
#         output_network_file <- paste(gsea_working_path,"screenshot_network.pdf",sep="/")
#         em_command2 = paste("view export OutputFile= ",output_network_file, " options=PDF", sep="")
#         commandsRun(em_command2)
#         em_command2 = paste("session save as file= ",paste(output_path,curr_model_name,sep="/"), sep="")
#         commandsRun(em_command2)
#       }
#     }
#   }, error=function(e){output_network_file=FALSE})
#   
#   }, silent=TRUE)
# }

```
`r if(gsea_section & species!="Other" & length(gene_data_index)>0 ){"Click below to view GSEA results:  "} `
`r if(gsea_section & species!="Other" & length(gene_data_index)>0){ output_links } `

`r if(enrichment_map ){if( class(output_network_file)!="logical"){ "Click to view EnrichmentMap screen shot: \n"}  }`
`r if(enrichment_map ){if( class(output_network_file)!="logical"){ paste(" | [ EnrichmentMap Image](",output_subdir,"/",gsea_working_dir,"/screenshot_network.pdf)", sep="") } } `  
`r if(enrichment_map ){if( class(output_network_file)!="logical"){ "If multiple contrasts, load GSEA results into Cytoscape/EnrichmentMap manually."} } `

---

```{r echo=TRUE}
run_ksea <- length(phos_data_index)>0 & grepl("Human", species)  & runDifferential;
```

`r if( run_ksea ){"### Kinase Enrichment"} `
```{r ksea_section, eval=(run_ksea)}
try({
output_links <- "";

dir_path <- file.path(output_path, "KSEA")
if( dir.exists(dir_path) == FALSE ) {dir.create(dir_path)}

for( i in 1:length(phos_data_index) ){
  
  for( c in 1:length(loop_list) ){
    PX2 <- FALSE
    try({
      top_sum <- topTable(omicsList[[ phos_data_index[i] ]][["fit"]], adjust="BH", n=Inf, sort.by='p', coef=c);
      PX2 <- top_sum[,c("Protein","Gene","Sequence.window")]
      colnames(PX2) <- c("Protein", "Gene", "Peptide")
      PX2[,"Residue.Both"] <- paste(top_sum[,"Amino.acid"],top_sum[,"Position"], sep="")
      PX2[,"p"] <- top_sum[,"P.Value"]
      PX2[,"FC"] <- 2^top_sum[,"logFC"]
    })
    if( class(PX2)=="logical" ){
      PX2 <- fData(omicsList[[ phos_data_index[i] ]][["eSet"]])[,c("Protein","Gene","Sequence.window")]
      colnames(PX2) <- c("Protein", "Gene", "Peptide")
      PX2[,"Residue.Both"] <- paste(fData(omicsList[[ phos_data_index[i] ]][["eSet"]])[,"Amino.acid"],
                                    fData(omicsList[[ phos_data_index[i] ]][["eSet"]])[,"Position"], sep="")
      PX2[,"p"] <- "NULL";
      contrast_grab <- grep( paste("logfc_", gsub("-","_",contrast_strings[c]), sep=""), colnames(fData(omicsList[[ phos_data_index[i] ]][["eSet"]]) ) )
      PX2[,"FC"] <- 2^(fData(omicsList[[ phos_data_index[i] ]][["eSet"]])[,contrast_grab] )    
    }
    
    dir_name <- paste("KSEA_", contrast_strings[c], sep="")
    dir_path <- file.path(output_path, "KSEA", dir_name)
    if( dir.exists(dir_path) == FALSE ) {dir.create(dir_path)}
    
    try({
    KSEA.Complete(KSData, PX2, NetworKIN=TRUE, NetworKIN.cutoff=5, m.cutoff=3, p.cutoff=0.05, output_dir=dir_path)
    
    output_links <- paste(output_links,paste("[ ",dir_name," ](",output_subdir,"/KSEA/",dir_name,"/KSEA Bar Plot.tiff)  |  ",sep=""),sep="" )
    })
  }
}
})

```
`r if(run_ksea){"Click below to view KSEA results:  "} `
`r if( run_ksea ){ output_links } `

---
```{r echo=FALSE}
show_metab_enrich <- grepl(paste("Human", "Mouse", "Fruit Fly", "Yeast", "E. coli", sep="|"),species)
```

`r if( (show_metab_enrich& length(gene_data_index)>0 & runDifferential) ){"### GSEA for Metabolomic Models"} `
```{r gsea_met_section, eval=(show_metab_enrich & length(gene_data_index)>0 & runDifferential) }

output_links <- "";
run_gsea <- gsea_section; # if GSEA has already been run, set to false and include results in directory

if( run_gsea & species!="Other" ) { try({

  gsea_working_dir <- "GSEA_MOMENTA"
  gsea_working_path <- file.path(output_path, gsea_working_dir)
  gsea_images_path <- file.path(gsea_working_path, "Images")
  gsea_absval_path <- file.path(gsea_working_path, "AbsoluteValueRank")
  if( dir.exists(gsea_working_path) == FALSE ) { dir.create(gsea_working_path) }
  if( dir.exists(gsea_images_path) == FALSE ) { dir.create(gsea_images_path) }
  if( dir.exists(gsea_absval_path) == FALSE ) { dir.create(gsea_absval_path) }
  
  geneset_lookup <-  read.delim(file.path(gsub("src", "data", notebook_dir),"matched","MatchedGeneSets.txt"))
  
  gmt_files <- as.vector(geneset_lookup[which(geneset_lookup[,"Species"]==species),"GeneSet"])
  gmt_names <- as.vector(geneset_lookup[which(geneset_lookup[,"Species"]==species),"Name"])
  
for(gmt_index in 1:length(gmt_files)){
    
  dest_gmt_file <-file.path(gsea_working_path, gmt_files[gmt_index])
  file.copy(from=file.path( gsub("src", "data", notebook_dir),"matched",gmt_files[gmt_index]), to= dest_gmt_file )  
  
    analysis_names <- vector("list", (length(loop_list)*length(gene_data_index)) ) ;
    counter<-1;                       
    for (j in 1:length(gene_data_index)) {
      for (k in 1:length(loop_list)) {
        analysis_names[[counter]] <- paste(omicsList[[ gene_data_index[j] ]][["dataType"]], "_", gsub("-", "_", contrast_strings[k]), sep="") ;
        counter <- counter+1;
      }
    }
    rnk_files <- vector("list", (length(loop_list)*length(gene_data_index)) ) ;
    counter<-1;                       
    for (j in 1:length(gene_data_index)) {
      for (k in 1:length(loop_list)) {
        if(grepl("Human", species)){ 
          rnk_files[[counter]] <- file.path(paste(output_contrast_path_files,"/GSEA_",omicsList[[ gene_data_index[j] ]][["dataType"]],
                                                  "_",gsub("-","_",contrast_strings[k]),".rnk", sep=""))
        } else {
          rnk_files[[counter]] <- file.path(paste(output_contrast_path_files,"/GSEA_",
                                                  omicsList[[ gene_data_index[j] ]][["dataType"]],"_Uppercase_",
                                                  gsub("-","_",contrast_strings[k]),".rnk", sep=""))
        }
        counter <- counter+1;
      }
    }
    expression_files <- vector("list", (length(loop_list)*length(gene_data_index)) ) ;
    counter<-1;                       
    for (j in 1:length(gene_data_index)) {
      for (k in 1:length(loop_list)) {
        expression_files[[counter]] <- file.path(paste(output_files_subdir, "/Expression_matrix_",
                                                       omicsList[[ gene_data_index[j] ]][["dataType"]],".txt", sep=""))
        counter <- counter+1;
      }
    } 
      if(length(gene_data_index)>1){ 
    for (i in 1:length(loop_list)) {
      analysis_names[[length(analysis_names)+1]] <- paste("Combined_",gsub("-", "_", contrast_strings[i]),sep="")
    }
    for (i in 1:length(loop_list)){
        if(grepl("Human", species)){ 
          rnk_files <- append(rnk_files, file.path(output_contrast_path_files,
                                                        paste("GSEA_combined_",gsub("-","_",contrast_strings[i]),".rnk", sep="") ))
        } else {
          rnk_files <- append(rnk_files, file.path(output_contrast_path_files,
                                                        paste("GSEA_combined_Uppercase_",gsub("-","_",contrast_strings[i]),".rnk", sep="") ))
        }
    }
    for (i in 1:num_contrasts){  
        expression_files <- append(expression_files, file.path(output_files_subdir,paste("Expression_matrix_",
                                                                                              omicsList[[ gene_data_index[1] ]][["dataType"]],
                                                                                              ".txt", sep="")))
    }
  }
  
  for (i in 1:length(analysis_names)){
    analysis_name <- paste(analysis_names[[i]], gmt_names[gmt_index], sep="_")
    rnk_file <- rnk_files[[i]]
    expression_file <- expression_files[[i]]
    
    ranked_features <- read.delim(rnk_file)
    file.copy(from=rnk_file, to=file.path(gsea_working_path,basename(rnk_file)) )
    ranked_vector <- ranked_features[,"rank"]
    names(ranked_vector) <- ranked_features[,"GeneName"]
    pathways_gmt <- gmtPathways(dest_gmt_file)
    suppressWarnings({
    fgsea_results <- fgsea(pathways=pathways_gmt, 
                                     stats=ranked_vector,
                                     minSize=15, 
                                     maxSize=500, 
                                     nperm=10000)
    })
    
    fgsea_out <- fgsea_results[,c(1,1,2,3)]
    colnames(fgsea_out) <- c("Term", "Description", "p.Val", "FDR")
    fgsea_out[,"Phenotype"] <- sign(fgsea_results[,"NES"])
    fgsea_out[,"Genes"] <- apply(fgsea_results[,"leadingEdge"], 1, function(x) paste(as.character(unlist(x["leadingEdge"])), collapse=","))
    fgsea_out[,"Genes"] <- apply(fgsea_out, 1, function(x) gsub(" ", "", x["Genes"]))
    fgsea_out[,"NES"] <- (fgsea_results[,"NES"])
    fgsea_out[,"ES"] <- (fgsea_results[,"ES"])
    fgsea_out[,"Gene_Hits"] <- apply(fgsea_results, 1, function(x) length(unlist(x["leadingEdge"])) )
    fgsea_out[,"Gene_Total"] <- fgsea_results[,"size"]
    output_filename <- file.path(output_path, gsea_working_dir, paste("fgsea_", analysis_name, ".txt", sep=""))
    write.table(fgsea_out, file=output_filename, sep="\t", row.names=FALSE, quote=FALSE);
    
    topPathways <- fgsea_results[order(pval, decreasing=F), pathway]
    output_filename <- file.path(gsea_images_path, paste("fgsea_", analysis_name, "_allPathways.pdf", sep=""))
    pdf(output_filename, width=4, height=3)
    for(pathway_index in 1:length(topPathways)){ try({
      print(plotEnrichment(pathways_gmt[[topPathways[pathway_index] ]], ranked_vector) +
              labs(title=topPathways[pathway_index]) + theme(plot.title=element_text(size=6)) )
    }, silent=TRUE) }
    dev.off()
    
    
    ranked_features <- read.delim(rnk_file)
    ranked_vector <- abs(ranked_features[,"rank"])
    names(ranked_vector) <- ranked_features[,"GeneName"]
    pathways_gmt <- gmtPathways(dest_gmt_file)
    suppressWarnings({
    fgsea_results <- fgsea(pathways=pathways_gmt, 
                                     stats=ranked_vector,
                                     minSize=15, 
                                     maxSize=500, 
                                     nperm=10000)
    })
    
    fgsea_out <- fgsea_results[,c(1,1,2,3)]
    colnames(fgsea_out) <- c("Term", "Description", "p.Val", "FDR")
    fgsea_out[,"Phenotype"] <- sign(fgsea_results[,"NES"])
    fgsea_out[,"Genes"] <- apply(fgsea_results[,"leadingEdge"], 1, function(x) paste(as.character(unlist(x["leadingEdge"])), collapse=","))
    fgsea_out[,"Genes"] <- apply(fgsea_out, 1, function(x) gsub(" ", "", x["Genes"]))
    fgsea_out[,"NES"] <- (fgsea_results[,"NES"])
    fgsea_out[,"ES"] <- (fgsea_results[,"ES"])
    fgsea_out[,"Gene_Hits"] <- apply(fgsea_results, 1, function(x) length(unlist(x["leadingEdge"])) )
    fgsea_out[,"Gene_Total"] <- fgsea_results[,"size"]
    output_filename <- file.path(gsea_absval_path, paste("fgsea_", analysis_name, "_AbsVal.txt", sep=""))
    write.table(fgsea_out, file=output_filename, sep="\t", row.names=FALSE, quote=FALSE);
    
  }

}

}) } 

```

---

`r if( length(metab_data_index)>0 ){"### Metabolite Enrichment"}`
```{r mummichog, eval=( length(metab_data_index)>0 )}

output_links <- "";
output_mummichog_subdir <- file.path(output_subdir,"Metabo_Enrichment" )
output_mummichog_path <- file.path(working_dir, output_mummichog_subdir)
if( dir.exists(output_mummichog_path) == FALSE ) { dir.create(output_mummichog_path) }

geneset_lookup <-  read.delim(file.path(gsub("src", "data", notebook_dir),"metabolite_model_table.txt"))
mumm_libs <- as.vector(geneset_lookup[which(geneset_lookup[,"Species"]==species),"MetLib"])

suppressWarnings({ suppressMessages({
for(i in 1:length(metab_data_index)){ try({
  for(mumm_lib_index in 1:length(mumm_libs)) {
    for (c in 1:length(loop_list)){ try({
      analysis_name <- paste(mumm_libs[mumm_lib_index],"_",omicsList[[ metab_data_index[i] ]][["dataType"]],
                             "_",gsub("-","_",contrast_strings[c]), sep="" )
      
      if(contrast_strings[1]!="Abundance"){ 
        output_mummichog_subdir_tmp <- file.path(output_mummichog_subdir, analysis_name)
        output_mummichog_path_tmp <- file.path(working_dir, output_mummichog_subdir_tmp)
        if( dir.exists(output_mummichog_path_tmp) == FALSE ) { dir.create(output_mummichog_path_tmp) }
        setwd(output_mummichog_path_tmp)
        
        input_data <- na.omit(topTable(omicsList[[ metab_data_index[i] ]][["fit"]],
                                       adjust="BH", n=5000, sort.by='p', coef=c)[,c("mz", "P.Value", "t")] );
        colnames(input_data) <- c("m.z", "p.value", "t.score")
        output_filename <- file.path(output_mummichog_path,paste("mummichog_",omicsList[[ metab_data_index[i] ]][["dataType"]],
                                                                 "_",gsub("-","_",contrast_strings[c]),".txt", sep=""));
        write.table(x=input_data, file=output_filename, sep='\t',row.names=FALSE, col.names=TRUE, quote=FALSE)
      
        if( grepl("Negative.Ion", omicsList[[ metab_data_index[i] ]][["dataFormat"]] ) ) { ion_mode<-"negative" } else { ion_mode<-"positive"}
      
        mSet <- InitDataObjects("mass_all", "mummichog", FALSE);
        #SetPeakFormat("mpt")
        mSet <- Read.PeakListData(mSet, output_filename);
        mSet <- MetaboAnalystR::UpdateMummichogParameters(mSet, "0.1", ion_mode, 0.05);
        mSet <- SanityCheckMummichogData(mSet);
      
        mSet <- PerformMummichog(mSet, mumm_libs[mumm_lib_index], permNum=1000);
        mSet <- PerformGSEA(mSet, mumm_libs[mumm_lib_index], permNum=1000);
      
        mSet <- PlotIntegPaths(mSet);
        #PreparePDFReport(mSet, usrName=analysis_name);
      
        tmp_table <- read.delim("mummichog_fgsea_pathway_enrichment.csv", sep=",", stringsAsFactors=FALSE)
        file.remove("mummichog_fgsea_pathway_enrichment.csv")
        tmp_table[,1] <- toupper(tmp_table[,1])
        out_table <- tmp_table[,c(1,1, 4, 5)]
        names(out_table) <- c("Term", "Description", "p.Val", "FDR")
        out_table[,"Phenotype"] <- "+1"
        out_table[tmp_table[,"NES"]<0,"Phenotype"] <- "-1"
        out_table[,"Gene_Hits"] <- tmp_table[,"Hits"]
        out_table[,"Gene_Total"] <- tmp_table[,"Pathway_Total"]
        out_table[,"Gene_Ratio"] <- out_table[,"Gene_Hits"]/out_table[,"Gene_Total"]
        out_table[,"NES"] <- tmp_table[,"NES"]
        out_table[,"Ionization"] <- ion_mode
        write.table(out_table,file="emap_fgsea_pathway_enrichment.txt", sep="\t", quote=FALSE, row.names=FALSE)
      
        if(nrow(out_table)>15){ out_table <- out_table[1:15,] }
        drawEnrichment(out_table,type=analysis_name,label="fgsea", outputpath=output_mummichog_path_tmp)
      
        tmp_table <- read.delim("mummichog_pathway_enrichment.csv", sep=",", stringsAsFactors=FALSE)
        file.remove("mummichog_pathway_enrichment.csv")
        tmp_table[,1] <- toupper(tmp_table[,1])
        out_table <- tmp_table[,c(1,1, 7)]
        names(out_table) <- c("Term", "Description", "p.Val")
        #out_table[,"Phenotype"] <- "+1"
        out_table[,"Gene_Hits"] <- tmp_table[,"Hits.total"]
        out_table[,"Gene_Total"] <- tmp_table[,"Pathway.total"]
        out_table[,"Gene_Ratio"] <- out_table[,"Gene_Hits"]/out_table[,"Gene_Total"]
        out_table[,"Ionization"] <- ion_mode
        write.table(out_table,file="emap_mummi_pathway_enrichment.txt", sep="\t", quote=FALSE, row.names=FALSE)
      
        #if(nrow(out_table)>15){ out_table <- out_table[1:15,] }
        #drawEnrichment(out_table,type=analysis_name,label="mummichog", outputpath=output_mummichog_path_tmp)
        #mset<-PlotIntegPaths(mSet, "png", 400, width=10)
      
        add_link <- paste("[ ",analysis_name, " ](", output_mummichog_subdir_tmp,"/emap_fgsea_pathway_enrichment.txt)", sep="")
        output_links <- paste(output_links, add_link, sep=" | " )
      
      } else { # Run based on enrichment of most intense/least intense
        input_data <- data.frame(m.z= fData(omicsList[[ metab_data_index[i] ]][["eSet"]])$mz,
                                 rank= rowMeans(exprs(omicsList[[ metab_data_index[i] ]][["eSet"]])) )
        input_data <- input_data[order(input_data[,"rank"], decreasing=T), ]
        output_filename <- file.path(output_mummichog_path,paste("mummichog_",omicsList[[ metab_data_index[i] ]][["dataType"]],
                                                                 "_",gsub("-","_",contrast_strings[c]),".txt", sep=""));
        write.table(x=input_data[,"m.z"], file=output_filename, sep='\t',row.names=FALSE, col.names=FALSE, quote=FALSE)
        
      }
    }) }
    output_links <- paste(output_links, add_link, sep="  \n" )
  }
}) }
}) })
    
setwd(working_dir)
    
```
`r if(( length(metab_data_index)>0 ) ){"Click below to view Mummichog results:  "}`
`r if(( length(metab_data_index)>0 ) ){ if(output_links != ""){ output_links }}`

---

```{r child=file.path(notebook_dir, 'Notebook_4_Integ.Rmd'), eval=( show_metab_enrich & length(metab_data_index>1))}
```


