# QC and Exploratory Analysis

`r if(txtFolder){"### Raw Data QC Report (PTXQC)"}`
```{r PTX_QC_report, results='hide'}
# if txt folder provided, generate PTXQC report
tryCatch(expr={
if(txtFolder){
  txt_folder <- file.path(working_dir, "txt");
  report.mq <- PTXQC::createReport(txt_folder);
}
}, error=function(e){print("Error generating PTXQC report")} )
```
`r if(txtFolder){"[ View Raw Data QC Report (PTXQC) ](txt/report_v0.92.3_MQ_pipeline.pdf)" }` 

---

### Data Normalization
```{r normalization, fig.width=12, fig.height=4}

g = g.norm(g)
g = g.norm.vis(g)

print_report <- "";
for(i in 1:length(g$omicsList)){
    print_report <- paste(print_report, 
                          paste(g$omicsList[[i]][[1]],": ",nrow(g$omicsList[[i]][[4]])," features detected in raw data, ", 
                                nrow(g$omicsList[[i]][[5]]), " after filtering. \n",sep=""),
                          paste(g$omicsList[[i]][[1]],": ", ncol(g$omicsList[[i]][[5]]), " samples after filtering. \n",sep=""), sep="" )
}

cat(print_report)
writeLines(print_report, con=file.path(g$output_files_path, "Feature_Numbers.txt"))

nulls=sapply(g$pca_output,is.null)
if(all(nulls)){
  cat("Warning: no PCA plots could be created for this data.")
} else {
  if(any(nulls))
    cat("Warning: some PCA plots could not be created for this data.")
  try({ gridExtra::grid.arrange(grobs=g$pca_output[!nulls], ncol=2) })
}

```
Click to view full size images:  
`r { g$output_links }`  

---

`r if(length(g$omicsList)>1){"### Venn Diagram"}`
```{r venn, eval= (length(g$omicsList)>1) }

g = g.venn(g)

```
`r if(length(g$omicsList)>1){"Click to view full size overlap plots: "}` `r if(length(g$omicsList)>1){ g$venn_output_links }`  

---

```{r drawcorplot, eval= (length(g$gene_data_index)>0 | length(g$prot_data_index)>1) }

g = g.corplot(g)

```

`r if(length(g$gene_data_index)>1 | length(g$prot_data_index)>1){"### Correlation Across Datasets"}`
```{r gene_correlation, eval= (length(g$gene_data_index)>1 | length(g$prot_data_index)>1) }

```
`r if(length(g$gene_data_index)>1 | length(g$prot_data_index)>1){"Click to view full size correlation plots: "}`
`r if(length(g$gene_data_index)>1 | length(g$prot_data_index)>1){ g$output_links }`  

---

### Correlation Across Groups
```{r sample_group_correlation, results='hide'}

g = g.cor.across.groups(g)

```
`r {"Click to view full size correlation plots: "}` `r { g$output_links }`  

---

`r if( length(g$prot_data_index)>0 | length(g$gene_data_index)>1 ){"### Normalize to 1st Data Set"}`
```{r site_normalization, eval= ( length(g$prot_data_index)>0 | length(g$gene_data_index)>1 ) }
suppressMessages({suppressWarnings({ 
  g = g.norm.to.first(g)
})})
```
`r if( length(g$prot_data_index)>0 | length(g$gene_data_index)>1 ){"Proteome Normalization:"}`
`r if( length(g$prot_data_index)>0 | length(g$gene_data_index)>1 ){ g$output_links }`

```{r add_site_norm, eval=use_site_norm}
g = g.use.site.norm(g)

```

---

### Variation Plots, Filtering
```{r variation}

g = g.varplot(g)
  
```
Click below to view SD vs Mean and Pairs plots: `r { g$output_links }`  

---

### Global Static Heatmaps
```{r heatmaps_qc, results='hide'}

if( exists("combined_met") && combined_met && length(g$mz_data_index) > 1 ){ # make combined eset for metabolomics
  g = g.combine.met(g)
}
g = g.heatmap.static(g)

```
Click to view static heatmaps plots: `r { g$output_links }`  

---

```{r boxplots_for_select_features}

# If there's a directory called "Subset_Lists" make feature lists based on text files
g = g.boxplots(g)

```

---

`r if(int_heatmap_section){"### Interactive Heatmap"}`
```{r interactive_plots_qc, eval=int_heatmap_section}
g = g.interactive.heatmap(g)

```
`r if(int_heatmap_section){ paste("Upload files to [Morpheus](https://software.broadinstitute.org/morpheus/) for viewing. Files are located at: ", g$output_links) }`  

---

### Save Data
```{r save_data_qc }

g = g.save.data.qc(g)

``` 

---

