# QC and Exploratory Analysis

`r if(txtFolder){"### Raw Data QC Report (PTXQC)"}`
```{r PTX_QC_report, results='hide'}
# if txt folder provided, generate PTXQC report
tryCatch(expr={
if(txtFolder){
  library(PTXQC)
  require(methods)
  library(data.table)

  txt_folder <- file.path(working_dir, "txt");
  report.mq <- createReport(txt_folder);
}
}, error=function(e){print("Error generating PTXQC report")} )
```
`r if(txtFolder){"[ View Raw Data QC Report (PTXQC) ](txt/report_v0.92.3_MQ_pipeline.pdf)" }` 

---

### Data Normalization
```{r normalization, fig.width=12, fig.height=4}

pca_output <- vector("list", length(omicsList) )
output_links<-"";

for(i in 1:length(omicsList)){
  if(length(zero_percent)==1){use_zero<-zero_percent}else{use_zero<-zero_percent[i]}
  if(length(norm_method)==1){use_norm<-norm_method}else{use_norm<-norm_method[i]}
  
  if(debug_opt){ index<-14 } else { index <- 5 }
  omicsList[[i]][[5]] <-intensityNorm(eset=omicsList[[i]][[index]],
                                      type=omicsList[[i]][[1]], 
                                      data_format=omicsList[[i]][["dataFormat"]],
                                      norm=norm_method, zero_cutoff=zero_percent,
                                      min_feature=min_feature_per_sample,
                                      norm_by_batch=norm_batches)
  
  
  # Make output hyperlinks for markdown
  add_link <- paste(omicsList[[i]][["dataType"]], ": [ Boxplot/Dist. ](", output_plots_subdir,
                    "/boxplot_histogram_",omicsList[[i]][["dataType"]],".pdf)", sep="")
  output_links <- paste(output_links, add_link, sep="" )
  
  # Batch correction
  if("Batch" %in% colnames(pData(omicsList[[i]][["eSet"]])) & !("Met" %in% omicsList[[i]][["dataFormat"]])) { try({
    suppressPackageStartupMessages({require(sva) });
    suppressMessages({ suppressWarnings({
    omicsList[[i]][[12]] <- omicsList[[i]][["eSet"]]
    names(omicsList[[i]])[12] <- "prebatch_eset"
    
    pheno = pData(omicsList[[i]][["eSet"]])
    edata = exprs(omicsList[[i]][["eSet"]])
    batch = pheno$Batch
  
    modcombat = model.matrix(~1, data=pheno)
    combat_edata = ComBat(dat=edata, batch=batch, mod=modcombat, par.prior=TRUE, prior.plots=FALSE)
    combat_eset = ExpressionSet(assayData=combat_edata)
    pData(combat_eset) = pData(omicsList[[i]][["eSet"]])
    fData(combat_eset) = fData(omicsList[[i]][["eSet"]])
    })})
    omicsList[[i]][["eSet"]] <- combat_eset
    tmp<-drawPCA(eset=omicsList[[i]][["prebatch_eset"]], type=paste(omicsList[[i]][["dataType"]],"_precorrection", sep=""));
    
    capture.output( intensityNorm(eset=omicsList[[i]][[index]],
                                      type=paste(omicsList[[i]][[1]],"_postBatchCor",sep=""), 
                                      data_format=omicsList[[i]][["dataFormat"]],
                                      norm=norm_post_batch, zero_cutoff=zero_percent,
                                      min_feature=min_feature_per_sample, norm_by_batch=F) )
    
  }) }
  
  if(remove_group!=""){
    omicsList[[i]][[5]] <- omicsList[[i]][[5]][, !grepl(paste(remove_group, collapse="|"),pData(omicsList[[i]][[5]])$Group) ]
  }
  
  try({
  # PCA plots
  pca_output[[i]] <- drawPCA(eset=omicsList[[i]][["eSet"]], type=omicsList[[i]][["dataType"]])
  
  # Make output hyperlinks for markdown
  add_link <- paste( "[ PCA ](", output_plots_subdir,"/PCAplots_",omicsList[[i]][["dataType"]],".pdf)", sep="")
  output_links <- paste(output_links, add_link, sep=" | " )
  })
  try({ 
    drawUMAP(eset=omicsList[[i]][["eSet"]], type=omicsList[[i]][["dataType"]])
    add_link <- paste( "[ UMAP ](", output_plots_subdir,"/UMAPplot_",omicsList[[i]][["dataType"]],".pdf)", sep="")
    output_links <- paste(output_links, add_link, sep=" | " )
  })
  
  output_links <- paste(output_links, "  \n", sep="" )
}

print_report <- "";
for(i in 1:length(omicsList)){
    print_report <- paste(print_report, 
                          paste(omicsList[[i]][[1]],": ",nrow(omicsList[[i]][[4]])," features detected in raw data, ", 
                                nrow(omicsList[[i]][[5]]), " after filtering. \n",sep=""),
                          paste(omicsList[[i]][[1]],": ", ncol(omicsList[[i]][[5]]), " samples after filtering. \n",sep=""), sep="" )
}

cat(print_report)
writeLines(print_report, con=file.path(output_files_path, "Feature_Numbers.txt"))

try({ grid.arrange(grobs=pca_output, ncol=2) })

```
Click to view full size images:  
`r { output_links }`  

---

`r if(length(omicsList)>1){"### Venn Diagram"}`
```{r venn, eval= (length(omicsList)>1) }

output_links <-"";

draw_vdata=function(data_index,var,name){
  v_list <- vector("list", length(data_index))
  for(i in 1:length(data_index)){
    v_list[[i]] <- fData(omicsList[[ data_index[i] ]][["eSet"]])[,var]
    if(class(v_list[[i]])=="numeric"){
      v_list[[i]] <- na.omit(round(v_list[[i]],digits=2))
    }
    names(v_list)[i] <- omicsList[[ data_index[i] ]][["dataType"]]
  }
  drawVenn(item_list=v_list, item_name=name)
  paste0("[ ",name," ](", output_plots_subdir,"/VennDiagram_",name,".pdf)")
}
output_links <- paste(c(
  if( length(gene_data_index) > 1 ) draw_vdata(gene_data_index,"Gene","Genes"),
  if( length(prot_data_index) > 1 ) draw_vdata(prot_data_index,"Protein","Proteins"),
  if( length(mz_data_index) > 1 ) draw_vdata(mz_data_index,"mz","Metabolites_mz")
  ), collapse=" | ")

```
`r if(length(omicsList)>1){"Click to view full size overlap plots: "}` `r if(length(omicsList)>1){ output_links }`  

---

```{r drawcorplot, eval= (length(gene_data_index)>0 | length(prot_data_index)>1) }

drawcorplot <- function(item_name,data_index,dataname=NULL){ try({
  output_links <-"";
  if(is.null(dataname)) dataname <- rep("eSet",length(data_index))
  fprefix=paste0(dataname[length(data_index)],"_")
  outname <- paste0(sub("eSet_","",fprefix),contrastgroups)

  for (j in 1:length(contrastgroups) ){
    group_name <- paste0(item_name,"_", outname[j])
    gene_values <- vector("list", (length(data_index)) )
    for (i in 1:(length(data_index)) ){
      group_columns <- grepl(contrastgroups[j], pData(omicsList[[ data_index[i] ]][[ dataname[i] ]])$Group)
      gene_values[[i]] <- data.frame( exprs(omicsList[[ data_index[i] ]][[ dataname[i] ]])[,group_columns] ); # Make a data frame of values and genes
      gene_values[[i]][,item_name] <- fData(omicsList[[ data_index[i] ]][[ dataname[i] ]])[,item_name] ;                                                                         
      gene_values[[i]] <- gene_values[[i]] %>% gather(sample, value, -all_of(item_name)); # Gather values and get average by gene
      gene_values[[i]] <- gene_values[[i]] %>% group_by(.dots=all_of(item_name)) %>% summarize(mean=mean(value));
      gene_values[[i]] <- gene_values[[i]][gene_values[[i]][,item_name]!="",]
      names(gene_values)[i] <- paste0(omicsList[[ data_index[i] ]][["dataType"]],"_",outname[j])
    }
    drawXYCorr(item_list=gene_values, item_name=item_name, file_name = outname[j] , subset_genes=subset_genes);
    add_link <- paste0("[ ",group_name," Correlation Plots ](",output_plots_subdir,"/Correlation_Plots_",group_name,".pdf)")
    output_links <- paste(output_links, add_link, sep=" | " )
  }
  output_links
}) }

```

`r if(length(gene_data_index)>1 | length(prot_data_index)>1){"### Correlation Across Datasets"}`
```{r gene_correlation, eval= (length(gene_data_index)>1 | length(prot_data_index)>1) }

# If there's a directory called "Subset_Lists" make feature lists based on text files
subset_genes=FALSE;
if( dir.exists("Subset_Lists") ){ try({
  filelist <- list.files(path=file.path("Subset_Lists"), pattern="*.txt")
  subset_genes <- vector("list", length=length(filelist))
  names(subset_genes) <- gsub("*.txt", "", filelist)
  
  for( i in 1:length(filelist)){
    subset_genes[i] <- unique(list(scan(file.path("Subset_Lists", filelist[i]), what="", sep=" " ) ))
  }
}) }

output_links <- paste(c(
  if(length(gene_data_index)>1) drawcorplot(item_name="Gene",data_index=gene_data_index),
  if(length(prot_data_index)>1) drawcorplot(item_name="Protein",data_index=prot_data_index)
  ), collapse=" | ")

```
`r if(length(gene_data_index)>1 | length(prot_data_index)>1){"Click to view full size correlation plots: "}`
`r if(length(gene_data_index)>1 | length(prot_data_index)>1){ output_links }`  

---

### Correlation Across Groups
```{r sample_group_correlation, results='hide'}

output_links <-"";

bplapply(1:length(omicsList),FUN=function(i){try({
  if(length(unique(pData(omicsList[[i]][["eSet"]])$Group))>1){
  gene_values <- data.frame(t(exprs(omicsList[[i]][["eSet"]])))
  gene_values$Group <- pData(omicsList[[i]][["eSet"]])$Group
  gene_values <- aggregate(. ~ Group, gene_values, mean)
  rownames(gene_values)<- gene_values$Group
  gene_values$Group <- NULL
  gene_values <- data.frame(t(gene_values))
  gene_values$Groups<- row.names(gene_values)
  
  item_list <- vector("list", ncol(gene_values)-1)
  for( j in 1:(ncol(gene_values)-1) ){
    item_list[[j]] <- as_tibble(cbind(gene_values[,"Groups"], gene_values[,j] ), .name_repair="minimal")
    names(item_list[[j]]) <-  c("Groups", colnames(gene_values)[j])
    item_list[[j]][,2] <- as.double(unlist(item_list[[j]][,2]))
  }
  names(item_list)<- colnames(gene_values)[1:(ncol(gene_values)-1)]
  file_name <- omicsList[[i]][["dataType"]]
  item_name <- "Groups"
  drawXYCorr(item_list=item_list, item_name=item_name, file_name=file_name);
  add_link <- paste("[ ",file_name," ](",output_plots_subdir,"/Correlation_Plots_",item_name,"_",file_name,".pdf)", sep="");
  output_links <- paste(output_links, add_link, sep=" | " )
  }
}  ) })

```
`r {"Click to view full size correlation plots: "}` `r { output_links }`  

---

`r if( length(prot_data_index)>0 | length(gene_data_index)>1 ){"### Normalize to 1st Data Set"}`
```{r site_normalization, eval= ( length(prot_data_index)>0 | length(gene_data_index)>1 ) }
suppressMessages({suppressWarnings({ 
output_links<-"";
try({
  
# Take first data set for normalization
if(length(prot_data_index)>1){
  dataname=c("eSet",rep("siteNorm",length(prot_data_index)-1))
  item_name="Protein"
  data_index=prot_data_index
} else {
  dataname=c("eSet",rep("siteNorm",length(gene_data_index)-1))
  item_name="Gene"
  data_index=gene_data_index
}

gene_values <- omicsList[[ data_index[1] ]][["eSet"]]; 
for(i in 2:length(data_index)){
  if( identical(gene_values$SampleName, omicsList[[ data_index[i] ]][["eSet"]]$SampleName) ) {
  
    norm_values <- omicsList[[ data_index[i] ]][["eSet"]] ;
    norm_values <- norm_values[ fData(norm_values)[,item_name] %in% fData(gene_values)[,item_name] ,];
 
    for(g in 1:nrow(norm_values)) {
      gene_values_rows <- grep(fData(norm_values)[g,item_name], fData(gene_values)[,item_name])
      if(length(gene_values_rows)>1){
        gene_values_means <- colMeans(exprs(gene_values)[gene_values_rows,])
      } else {
        gene_values_means <- (exprs(gene_values)[gene_values_rows,])
      }
      #XXX TODO +2*mean? (prot 1x, gene 2x ????)
      exprs(norm_values)[g,] <- (as.numeric(exprs(norm_values)[g,] - gene_values_means )) + mean(gene_values_means) ;
    }
  
    omicsList[[ data_index[i] ]][[10]] <- norm_values;
    names( omicsList[[ data_index[i] ]] )[10] <- "siteNorm";

    data_norm_index <- c(data_norm_index, data_index[i])
  }
}

if(length(prot_data_index)>1){
  drawcorplot(item_name="Protein",data_index=prot_data_index,dataname=dataname)
} else {
  drawcorplot(item_name="Gene",data_index=gene_data_index,dataname=dataname)
}

for(i in 1:length(data_norm_index)){
  type_name <- paste0(omicsList[[ data_norm_index[i] ]][["dataType"]], "_NormTo",omicsList[[ data_index[1] ]][["dataType"]])
  tmp<-drawPCA(eset=omicsList[[ data_norm_index[i] ]][["siteNorm"]], type=type_name, show_sample_names=TRUE)

  add_link <- paste0("[ PCA:",type_name, " ](", output_plots_subdir,"/PCAplots_",type_name,".pdf)" )
  output_links <- paste(output_links, add_link, sep=" | " )
}
  
})
})})
```
`r if( length(prot_data_index)>0 | length(gene_data_index)>1 ){"Proteome Normalization:"}`
`r if( length(prot_data_index)>0 | length(gene_data_index)>1 ){ output_links }`

```{r add_site_norm, eval=use_site_norm}

unnorm_gene_index <- gene_data_index
gene_data_index <- data_index[1]

for(i in 1:length(data_norm_index)){try({
  start_length <- length(omicsList) + 1
  omicsList[[start_length]] <- omicsList[[ data_norm_index[i] ]]
  omicsList[[start_length]][["eSet"]] <- omicsList[[start_length]][["siteNorm"]] 
  omicsList[[start_length]][["dataType"]] <- paste(omicsList[[start_length]][["dataType"]],
                                                   "_NormTo",omicsList[[ data_index[1] ]][["dataType"]],sep="")
  gene_data_index <- c(gene_data_index, start_length)
}) }

```

---

### Variation Plots, Filtering
```{r variation}
output_links<-"";

# Pairs, Variation, and Corrplot
for(i in 1:length(omicsList)){ try({
  omicsList[[i]][[6]] <- variationPlot(eset=omicsList[[i]][["eSet"]], type=omicsList[[i]][["dataType"]]);
  names(omicsList[[i]])[6] <- "topVariable";
    
  # Make output hyperlinks for markdown
  type_name <- omicsList[[i]][["dataType"]]
  add_link <- paste("[ ",type_name, "-Variation ](", output_plots_subdir,"/variation_",type_name,".pdf) | ",
                    "[ ",type_name, "-Correlation ](", output_plots_subdir,"/corrplot_",type_name,".pdf) | ",
                    "[ ",type_name, "-Pairs ](",output_plots_subdir,"/pairs_",type_name,".jpeg)",
                    sep="");
  
  output_links <- paste(output_links, add_link, sep="  \n" );
}) }

# Coefficient of Variation
try({
CVs <- data.frame( CV=apply( exprs(omicsList[[1]][["eSet"]]), 1, function(x) { (sd(x)/mean(x))*100 } ) )
CVs$Dataset <- omicsList[[1]][["dataType"]]
if( length(omicsList)>1 ){ for(i in 2:length(omicsList) ){
  tmp <- data.frame( CV=apply( exprs(omicsList[[i]][["eSet"]]), 1, function(x) { (sd(x)/mean(x))*100 } ) )
  tmp$Dataset <- omicsList[[i]][["dataType"]]
  CVs <- rbind(CVs, tmp)
}}

plot <- ggplot(CVs, aes(x=CV, fill=Dataset)) + geom_density(alpha=0.2)+ theme_bw() + 
    labs(title=paste("Coefficient of Variation Plot \n ", sep=''),
         x="Coefficient of Variation (%)", y="Frequency");

output_filename<-file.path(output_plots_path, paste("CV_Plot",".pdf", sep=""))
pdf(output_filename, width=3, height=3)
print(plot+theme(legend.position="none"))
suppressWarnings(print(plot+ scale_x_continuous(limits=c(0, 20))+theme(legend.position="none")) )
suppressWarnings(print(plot+ scale_x_continuous(limits=c(0, 10))+theme(legend.position="none")) )
grid.arrange(g_legend(plot))
dev.off()

avg <- aggregate( CVs[,"CV"], list(CVs[,"Dataset"]), mean) 
colnames(avg) <- c("Dataset", "Average CV")
print(avg)
write.table(avg, file=file.path(output_files_path, "Average_CV.txt"), quote=F, sep="\t")

add_link <- paste("[ CV Plot ](", output_plots_subdir,"/CV_Plot",".pdf) ", sep="");
output_links <- paste(output_links, add_link, sep="  \n" );
})
  
```
Click below to view SD vs Mean and Pairs plots: `r { output_links }`  

---

### Global Static Heatmaps
```{r heatmaps_qc, results='hide'}
output_links<-"";

if( exists("combined_met") && combined_met && length(mz_data_index) > 1 ){ # make combined eset for metabolomics
  merge_eset=function(xl,yl){
    x=xl$eSet; y=yl$eSet
    rn=c(paste(xl$dataType,rownames(exprs(x))),paste(yl$dataType,rownames(exprs(y))))
    rd=rbind(exprs(x),exprs(y))
    rownames(rd)=rn
    ret=ExpressionSet(assayData=rd)
    pData(ret)=pData(x)
    inds=intersect(colnames(fData(x)),colnames(fData(y)))
    fr=rbind(fData(x)[,inds],fData(y)[,inds])
    rownames(fr)=rn
    fData(ret)=fr
    ret
  }
  newi=length(omicsList)+1
  mz_data_index=c(mz_data_index,newi)
  metab_data_index=c(metab_data_index,newi)
  omicsList[[newi]]=list(eSet=merge_eset(omicsList[[mz_data_index[1]]],omicsList[[mz_data_index[2]]]),dataType="met_combined",dataFormat="Metabolomics (combined)")
  omicsList[[newi]][["topVariable"]]=variationPlot(eset=omicsList[[newi]][["eSet"]], type=omicsList[[newi]][["dataType"]]);
}

# If there's a directory called "Subset_Lists" make feature lists based on text files
subset_genes=FALSE;
if( dir.exists("Subset_Lists") ){ try({
  filelist <- list.files(path=file.path("Subset_Lists"), pattern="*.txt")
  subset_genes <- vector("list", length=length(filelist))
  names(subset_genes) <- gsub("*.txt", "", filelist)
  
  for( i in 1:length(filelist)){
    subset_genes[i] <- unique(list(scan(file.path("Subset_Lists", filelist[i]), what="", sep=" " ) ))
  }
}) }

# Heatmaps
bplapply(1:length(omicsList),FUN=function(i){ try({
  subset_rows=FALSE;
  if( class(subset_genes)!="logical" ){ try({
    subset_rows <- vector("list", length(subset_genes))
    names(subset_rows) <- names(subset_genes)
    for( j in 1:length(subset_genes) ){
      if( "Gene" %in% colnames(fData(omicsList[[i]][["eSet"]])) ){
        subset_rows[[j]] <- rownames(omicsList[[i]][["eSet"]])[which(fData(omicsList[[i]][["eSet"]])$Gene %in% subset_genes[[j]] )]
      }
      if( length(subset_rows[[j]])==0 ){
        subset_rows[[j]] <- rownames(omicsList[[i]][["eSet"]])[ which(rownames(omicsList[[i]][["eSet"]]) %in% subset_genes[[j]] ) ]
      }
    }
  },silent=T) }
  # Draw the heatmaps and sace the Complex Heatmap object
  map_out <- drawHeatmaps(eset=omicsList[[i]][["eSet"]], emat_top=omicsList[[i]][["topVariable"]],
                          type=omicsList[[i]][["dataType"]], subset=subset_rows, k_clust=knn_heatmap);
  saveRDS(map_out, file=file.path(output_files_path, paste("Heatmap_",omicsList[[i]][["dataType"]],".RDS",sep="")) )
  
}) })

for(i in 1:length(omicsList)){
  # Make output hyperlinks for markdown
  add_link <- paste("[ ",omicsList[[i]][["dataType"]], " ](", output_plots_subdir,"/heatmaps_",omicsList[[i]][["dataType"]],".pdf)", sep="")
  output_links <- paste(output_links, add_link, sep=" | " )
}

```
Click to view static heatmaps plots: `r { output_links }`  

---

```{r boxplots_for_select_features}

# If there's a directory called "Subset_Lists" make feature lists based on text files
subset_genes=FALSE;
if( dir.exists("Subset_Lists") ){ try({
  filelist <- list.files(path=file.path("Subset_Lists"), pattern="*.txt")
  subset_genes <- vector("list", length=length(filelist))
  names(subset_genes) <- gsub("*.txt", "", filelist)
  
  for( i in 1:length(filelist)){
    subset_genes[i] <- unique(list(scan(file.path("Subset_Lists", filelist[i]), what="", sep=" " ) ))
  }
  subset_genes <- unique(unlist(subset_genes, use.names = F))
}) }

for(i in 1:length(omicsList)){ try({
  subset_rows=FALSE;
  if( class(subset_genes)!="logical" ){ try({
    subset_rows <- c()
    if( "Gene" %in% colnames(fData(omicsList[[i]][["eSet"]])) ){
      subset_rows <- rownames(omicsList[[i]][["eSet"]])[which(fData(omicsList[[i]][["eSet"]])$Gene %in% subset_genes )]
    }
    if( length(subset_rows)==0 ){
      subset_rows <- rownames(omicsList[[i]][["eSet"]])[ which(rownames(omicsList[[i]][["eSet"]]) %in% subset_genes ) ]
    }
  },silent=T) }
  # Draw the boxplots
  if(class(subset_rows)!="logical"){ 
    output_filename<-file.path(output_plots_path, paste(omicsList[[i]][["dataType"]],"_SelectFeatures",".pdf", sep=""))
    pdf(output_filename, width=3, height=3)
    for(j in 1:length(subset_rows)){ try({
      dat <- data.frame(Intensity=exprs(omicsList[[i]][["eSet"]])[subset_rows[j],],
                        Group = pData(omicsList[[i]][["eSet"]])$Group  )
      print(ggplot(data=dat, aes(color=Group, x=Group, y=Intensity)) + geom_boxplot() + geom_point() +
          theme_bw() + labs(title=paste( omicsList[[i]][["dataType"]],"\n", subset_rows[j], sep="")) +
          theme(legend.position = "none") + ylab("Log2 Intensity") +
          theme(axis.text.x=element_text(angle=45, hjust=1))
      )
    
    }) }
    dev.off()
  }
}) }

```

---

`r if(int_heatmap_section){"### Interactive Heatmap"}`
```{r interactive_plots_qc, eval=int_heatmap_section}
output_links<-"";

for(i in 1:length(omicsList)){
  # Make the interactive heatmap
  interactiveHeatmap(eset=omicsList[[i]][["eSet"]], type=omicsList[[i]][["dataType"]]);
  
  # Make output hyperlinks for markdown
  add_link <- paste("[ ",omicsList[[i]][["dataType"]], "-All ](", output_plots_subdir,"/heatmap_all_",
                    omicsList[[i]][["dataType"]],".html)", sep="");
  output_links <- paste(output_links, add_link, sep=" | " );
}

```
`r if(int_heatmap_section){"Click to view interactive heatmaps:"}`  `r if(int_heatmap_section){ output_links }`  

---

### Save Data
```{r save_data_qc }

# save for omics integrator, expression matrix, expression set RDS, ranked list for GSEA
subset_genes=FALSE;
if( dir.exists("Subset_Lists") ){ try({
  filelist <- list.files(path=file.path("Subset_Lists"), pattern="*.txt")
  subset_genes <- vector("list", length=length(filelist))
  names(subset_genes) <- gsub("*.txt", "", filelist)
  
  for( i in 1:length(filelist)){
    subset_genes[i] <- unique(list(scan(file.path("Subset_Lists", filelist[i]), what="", sep=" " ) ))
    subset_genes[[i]] <- subset_genes[[i]][subset_genes[[i]]!=""]
  }
}) }

for(i in 1:length(omicsList)){
  subset_rows=FALSE;
  if( class(subset_genes)!="logical" ){ try({
    subset_rows <- vector("list", length(subset_genes))
    names(subset_rows) <- names(subset_genes)
    for( j in 1:length(subset_genes) ){
      if( "Gene" %in% colnames(fData(omicsList[[i]][["eSet"]])) ){
        subset_rows[[j]] <- rownames(omicsList[[i]][["eSet"]])[which(fData(omicsList[[i]][["eSet"]])$Gene %in% subset_genes[[j]] )]
      }
      if( length(subset_rows[[j]])==0 ){
        subset_rows[[j]] <- rownames(omicsList[[i]][["eSet"]])[ grep( paste(subset_genes[[j]], collapse="|"), 
                                                                      rownames(exprs(omicsList[[i]][["eSet"]])) ) ]
      }
    }
  }, silent=TRUE) }
  saveFiles(data=omicsList[[i]], type=omicsList[[i]][["dataType"]], subset=subset_rows);
}  

# Save excel file summary for collaborators
if(saveXlsx==TRUE){    
  wbOut <- createWorkbook()
  
  for(i in 1:length(omicsList)){
    eSet=omicsList[[i]][["eSet"]][,order(pData(omicsList[[i]][["eSet"]])$Group)]
    writeDataToSheets(wb=wbOut, eset=eSet, type=omicsList[[i]][["dataType"]], data_format=omicsList[[i]][["dataFormat"]]);
  } 
  
  output_filename=file.path(output_path, paste(gsub("\\.","",make.names(project_name)), "_Summary", ".xlsx", sep=""))
  saveWorkbook(wbOut, file=output_filename, overwrite=TRUE)
}

``` 

---

