# Differential Analysis  

### Limma
```{r }
#limma
for(i in 1:length(omicsList)){

  # Create a model matrix and lmFit based on group - For more complex models, edit the following code in this chunk.
  f <- factor(pData(omicsList[[i]][["eSet"]])$Group)#, levels=contrastgroups); 
  #paste(pData(eset)Group, pData(eset1)$Disease, sep='.'))
  design <- model.matrix(~ 0 + f);
  colnames(design) <- make.names(levels(f));
  fit1 <- lmFit( omicsList[[i]][["eSet"]], design);
  
  # Make all contrasts between groups
  contrast_strings <- "";
  for (j in 1:(length(contrastgroups)-1)){
    for (k in (j+1):length(contrastgroups)){
      contrast_strings <- c(contrast_strings, paste(contrastgroups[k],'-',contrastgroups[j], sep=''));
  }  }
  contrast_strings <- contrast_strings[-1]
  contrast_matrix <- makeContrasts(contrasts=contrast_strings, levels=design);
    
  # Add the fit object to the omicsList
  fit1 <- contrasts.fit(fit1, contrast_matrix);
  omicsList[[i]][[7]] <- eBayes(fit1);
  names(omicsList[[i]])[7] <- "fit";

  # Summary of the fit
  end_col <- ncol(topTable(omicsList[[i]][["fit"]], adjust="BH"))
  start_col <- end_col - (3+length(contrast_strings))
  print(topTable(omicsList[[i]][["fit"]], adjust="BH")[,seq(start_col, end_col)])
  
  # Save the decide tests results and print summary 
  omicsList[[i]][[10]] <- decideTests(omicsList[[i]][["fit"]]);
  names(omicsList[[i]])[10] <- "dt";
  print(summary(omicsList[[i]][["dt"]]))
  #vennDiagram(omicsList[[i]][["dt"]]);

  # Save the information about the number of coeficients and the contrasts names
  omicsList[[i]][[8]] <- length(contrast_strings)
  names(omicsList[[i]])[8] <- "numCoef"
  omicsList[[i]][[9]] <- contrast_strings
  names(omicsList[[i]])[9] <- "contrasts"
  
}

# Make differential directory
output_contrast_subdir <- paste(output_subdir,"Differential", sep="/" )
output_contrast_path <- file.path(working_dir, output_contrast_subdir)
if( dir.exists(output_contrast_path) == FALSE ) { dir.create(output_contrast_path) }

```

---

### Volcano plots
```{r, message=FALSE, fig.width=4, fig.height=3}
output_links <- "";

for(i in 1:length(omicsList)){
  for(c in 1:omicsList[[i]][["numCoef"]]){
    # For the data set and coefficient, make the summary table and name
    top_sum <- topTable(omicsList[[i]][["fit"]], adjust="BH", n=Inf, sort.by='p', coef=c);
    type_name <- paste(omicsList[[i]][["dataType"]],omicsList[[i]][["contrasts"]][c], sep="_");
    
    # run Volcan function
    makeVolcano(top_sum, type=type_name);
    
    # Make output hyperlinks for markdown
    add_link <- paste("[ ",type_name, " ](", output_contrast_subdir,"/volcano_",type_name,".pdf)", sep="")
    output_links <- paste(output_links, add_link, sep=" | " )   
} }

```
Click to view static volcano plots: `r if(output_links != ""){ output_links }`  

---

### Differential Static Heatmaps
```{r, message=FALSE}
output_links <- "";

for(i in 1:length(omicsList)){
  for(c in 1:omicsList[[i]][["numCoef"]]){
    # For the data set and coefficient, make the summary table and name
    top_sum <- topTable(omicsList[[i]][["fit"]], adjust="BH", n=Inf, sort.by='p', coef=c, p.value=adjpcutoff);
    if (nrow(top_sum)<100){ top_sum <- topTable(omicsList[[i]][["fit"]], adjust="BH", n=100, sort.by='p', coef=c); }
    type_name <- paste(omicsList[[i]][["dataType"]],omicsList[[i]][["contrasts"]][c], sep="_");
    
    # run heatmap function
    drawthemapsDiff(eset=omicsList[[i]][["eSet"]], limmaSig=top_sum, type=type_name);
    
    # Make output hyperlinks for markdown
    add_link <- paste("[ ",type_name, " ](", output_contrast_subdir,"/heatmaps_",type_name,".pdf)", sep="");
    output_links <- paste(output_links, add_link, sep=" | " ); 
} }


```
Click to view static heatmap plots: `r if(output_links != ""){ output_links }`   


---

### Interactive Volcano Plots
```{r, warning=FALSE, message=FALSE}
output_links <- "";

for(i in 1:length(omicsList)){
  for(c in 1:omicsList[[i]][["numCoef"]]){
    top_sum <- topTable(omicsList[[i]][["fit"]], adjust="BH", n=Inf, sort.by='p', coef=c);
    type_name <- paste(omicsList[[i]][["dataType"]],omicsList[[i]][["contrasts"]][c], sep="_");

    makeInteractiveV(eset=omicsList[[i]][["eSet"]], fit=omicsList[[i]][["fit"]], dt=omicsList[[i]][["dt"]],
                     limmaSig=top_sum, type=type_name, col=c);

    add_link <- paste("[ ",type_name, " ](", output_contrast_subdir,"/InteractivePlots/Volcano-Plot_",type_name,".html)", sep="");
    output_links <- paste(output_links, add_link, sep=" | " ); 
}  }


```
Click to view interactive Volcano plots: `r if(output_links != ""){ output_links }`   


---

## Enrichment
`r if(enrichr_section){"### EnrichR"}`
```{r, message=FALSE, results='hide' }
# enrichR, save GSEA ranked list, eventually add run GSEA

output_links <- "";
if(enrichr_section==TRUE){
  
  # Run Enrichr for each data type
  for(i in 1:length(omicsList)){
    for(c in 1:omicsList[[i]][["numCoef"]]){
      top_sum <- topTable(omicsList[[i]][["fit"]], adjust="BH", n=Inf, sort.by='p', coef=c, p.value=adjpcutoff);
      if (nrow(top_sum)<50){ top_sum <- topTable(omicsList[[i]][["fit"]], adjust="BH", n=50, sort.by='p', coef=c); }
      type_name <- paste(omicsList[[i]][["dataType"]],omicsList[[i]][["contrasts"]][c], sep="_");
      
      runEnrichR(top_sum[, c("Gene", "logFC") ], type=type_name) 

      # Make output hyperlinks for markdown
      add_link <- paste("[ ",type_name, "-Up ](", output_contrast_subdir,"/enrichr_results_",type_name,"_up.txt) | ",
                      "[ ",type_name, "-Down ](", output_contrast_subdir,"/enrichr_results_",type_name,"_down.txt)", sep="")
      output_links <- paste(output_links, add_link, sep="  \n" )
  }  }
  
  # Run enrichr for combined data
  if(length(omicsList)>1){
    for(c in 1:omicsList[[i]][["numCoef"]]){
      type_name <- paste("Combined",omicsList[[i]][["contrasts"]][c], sep="_");
      
      gene_table <- topTable(omicsList[[1]][["fit"]], adjust="BH", n=Inf, sort.by='p', coef=c, p.value=adjpcutoff);
      if (nrow(gene_table)<50){ gene_table<-topTable(omicsList[[1]][["fit"]], adjust="BH", n=50, sort.by='p', coef=c)[, c("Gene", "logFC") ]; }
      for(i in 2:length(omicsList)){
        add_to_table <- topTable(omicsList[[i]][["fit"]], adjust="BH", n=Inf, sort.by='p', coef=c, p.value=adjpcutoff);
        if (nrow(add_to_table)<50){ add_to_table<-topTable(omicsList[[i]][["fit"]], adjust="BH", n=50, sort.by='p', coef=c); }
        gene_table <- rbind(gene_table, add_to_table[, c("Gene", "logFC") ])
      }

      runEnrichR(gene_table, type=type_name )
      
      # Make output hyperlinks for markdown
      add_link <- paste("[ ",type_name, "-Up ](", output_contrast_subdir,"/enrichr_results_",type_name,"_up.txt) | ",
                      "[ ",type_name, "-Down ](", output_contrast_subdir,"/enrichr_results_",type_name,"_down.txt)", sep="")
      output_links <- paste(output_links, add_link, sep="  \n" )
    }
  }
}

```
`r if(enrichr_section){"Click below to view enrichR results:  "}`
`r if(enrichr_section){ if(output_links != ""){ output_links }} `   

---

### Save Data
```{r, message=FALSE }
# save for omics integrator, expression matrix, expression set RDS, ranked list for GSEA

for(i in 1:length(omicsList)){
  for (c in 1:(omicsList[[1]][["numCoef"]])){
    top_sum <- topTable(omicsList[[i]][["fit"]], adjust="BH", n=Inf, sort.by='p', coef=c);
    type_name <- paste(omicsList[[i]][["dataType"]],gsub("-","_",omicsList[[i]][["contrasts"]][c]), sep="_");
    
    saveTheFilesDiff(omicsList[[i]][["eSet"]], top_sum, type=type_name, contrast_name=omicsList[[i]][["conrasts"]][c])
  }
}

if(length(omicsList)>1){
  for (c in 1:(omicsList[[1]][["numCoef"]])){
    top_sum <- topTable(omicsList[[1]][["fit"]], adjust.method="BH", n=Inf, sort.by='p', coef=c)
    ranked <- cbind(top_sum[,"Gene"], sign(top_sum[,"logFC"]) * -log10(top_sum[,"adj.P.Val"])) 

    for(i in 2:length(omicsList)){
      top_sum <- topTable(omicsList[[i]][["fit"]], coef=c, adjust.method="BH", n=Inf, sort.by='p')
      ranked1 <- cbind(top_sum[,"Gene"], sign(top_sum[,"logFC"]) * -log10(top_sum[,"adj.P.Val"])) 
      ranked <- rbind(ranked, ranked1)
    }
  
    colnames(ranked)<-c("GeneName", "rank")
    ranked <- ranked[ranked[,"GeneName"]!="", ]
    ranked <- ranked[order(abs(as.numeric(ranked[,"rank"])), decreasing=TRUE),]
    ranked <- ranked[!duplicated(ranked[,"GeneName"]),]
    ranked <- ranked[order(as.numeric(ranked[,"rank"]), decreasing=TRUE),]
    output_filename <- file.path(output_contrast_path,paste("GSEA_combined",gsub("-","_",omicsList[[i]][["contrasts"]][c]),".rnk", sep=""));
    write.table(x=ranked, file=output_filename, sep='\t',row.names=FALSE, col.names=TRUE, quote=FALSE)
  }
}
  
if(runDifferential==TRUE && saveXlsx==TRUE){   
  
  # Save excel file summary for collaborators
  wbOut <- createWorkbook()

  for(i in 1:length(omicsList)){
    writeDataToSheetsDiff(wb=wbOut, eset=omicsList[[i]][["eSet"]], limmaFit=omicsList[[i]][["fit"]], 
                          type=omicsList[[i]][["dataType"]], data_format=omicsList[[i]][["dataFormat"]]); 
  }  
  
  output_filename=file.path(output_path, paste(gsub("\\.","",make.names(project_name)), "_Summary", ".xlsx", sep=""))
  saveWorkbook(wbOut, file=output_filename, overwrite=TRUE)
  
}
```

---

`r if(gsea_section){"### GSEA and EnrichmentMap"}`
```{r, results='hide', message=FALSE, warning=FALSE, error=FALSE}
output_network_file <- FALSE;
output_links <- "";

if(gsea_section==TRUE) {
  tryCatch(expr = { library(RCurl)}, 
         error = function(e) { install.packages("RCurl")}, finally = library(RCurl))
  tryCatch(expr = { library(httr)}, 
         error = function(e) { install.packages("httr")}, finally = library(httr))
  tryCatch(expr = { library(RJSONIO)}, 
         error = function(e) { install.packages("RJSONIO")}, finally = library(RJSONIO))
  
  gsea_jar <- "/project/cnsbomic/Tools/gsea-3.0.jar"

  analysis_names <- vector("list", omicsList[[1]][["numCoef"]])
  for (i in 1:length(analysis_names)){
    analysis_names[[i]] <- gsub("-", "_", omicsList[[1]][["contrasts"]][i])
  }

  gsea_working_dir <- "GSEA"#gsub("\\.", "_", paste("GSEA_", analysis_name, sep=""))
  gsea_working_path <- file.path(output_path, gsea_working_dir)
  if( dir.exists(gsea_working_path) == FALSE ) { dir.create(gsea_working_path) }
  
  rnk_files <-vector("list", omicsList[[1]][["numCoef"]])
  for (i in 1:length(analysis_names)){  
    if(length(omicsList)>1){ 
      rnk_files[[i]] <- paste(output_contrast_path, "/GSEA_combined",gsub("-","_",omicsList[[1]][["contrasts"]][c]),".rnk", sep="")
    } else { rnk_files[[i]] <- paste(output_contrast_path,"/GSEA_",gsub("-","_",omicsList[[1]][["dataType"]]),"_",gsub("-","_",omicsList[[1]][["contrasts"]][c]),".rnk", sep="") }
  }
  expression_file <- file.path(paste(output_files_subdir, "/Expression_matrix_",omicsList[[1]][["dataType"]],".txt", sep=""))

  # Only if you need a new GMT file
  gmt_url = "http://download.baderlab.org/EM_Genesets/current_release/Human/symbol/"
  filenames = getURL(gmt_url)   #list all the files on the server
  tc = textConnection(filenames)
  contents = readLines(tc)
  close(tc)
  #get the gmt that has all the pathways and does not include terms inferred from electronic annotations(IEA), start with gmt file that has pathways only
  rx = gregexpr("(?<=<a href=\")(.*.GOBP_AllPathways_no_GO_iea.*.)(.gmt)(?=\">)",
                contents, perl = TRUE)
  gmt_file = unlist(regmatches(contents, rx))
  dest_gmt_file <- paste(gsea_working_path,paste("Supplementary_Table3_",gmt_file,sep="") ,sep="/")
  download.file(paste(gmt_url,gmt_file,sep=""),destfile=dest_gmt_file)

  if(run_gsea){
    for (i in 1:length(analysis_names)){
      analysis_name <- analysis_names[[i]]
      rnk_file <- rnk_files[[i]]
      command <- paste("java  -Xmx1G -cp",gsea_jar,  "xtools.gsea.GseaPreranked -gmx",
                       dest_gmt_file, "-rnk" , rnk_file,
                       #"-cls",paste(working.dir,cls_filename,sep="/"),
                       "-collapse false -nperm 1000 -permute gene_set -scoring_scheme weighted -rpt_label",
                       analysis_name,"  -num 100 -plot_top_x 20 -rnd_seed 12345  -set_max 200 -set_min 15", 
                       "-zip_report false -out" ,gsea_working_path, "-gui false > ",
                       paste(gsea_working_path,"gsea_output.txt", sep="/"),sep=" ")
      system(command)
      tmp<-list.files(working_dir) #clean up default output folder
      tmp<-tmp[file.info(tmp)$isdir==TRUE]
      tmp<-tmp[!is.na(tmp)]
      if(length(tmp)==1){
        if(nchar(tmp[1])==5){file.remove(paste(working_dir, tmp[1], sep="/")) }
      } 
    }
  }
  
  gsea_directories <- list.files(path = gsea_working_path, pattern = "\\.GseaPreranked")
  details = file.info(file.path(gsea_working_path, gsea_directories)) #get the details on the files
  details = details[with(details, order(as.POSIXct(mtime),decreasing = TRUE)), ]
  gsea_subdir_names <- gsub("(.*)/","", rev(row.names(details)[1:length(analysis_names)]))
  
  for (i in 1:length(analysis_names)){
    add_link <- paste("[ ",analysis_names[[i]], " ](", output_subdir,"/",gsea_working_dir,"/",gsea_subdir_names[i],"/index.html)", sep="");
    output_links <- paste(output_links, add_link, sep=" | " ); 
  }
  
  gsea_output_paths <- rev(row.names(details)[1:length(analysis_names)])
}

if(enrichment_map && gsea_section){
  #use easy cyRest library to communicate with cytoscape.
  tryCatch(expr = { library(devtools)}, 
         error = function(e) { install.packages("devtools")}, finally = library(devtools))
  tryCatch(expr = { library(r2cytoscape)}, 
          error = function(e) { install_github('cytoscape/cytoscape-automation/for-scripters/R/r2cytoscape')},
          finally = library(r2cytoscape))
  tryCatch(expr = { library(EasycyRest)}, 
          error = function(e) { install_github('BaderLab/Easycyrest/EasycyRest')},
          finally = library(EasycyRest))

  # Basic settings
  port.number = 1234
  base.url = paste("http://localhost:", toString(port.number), "/v1", sep="")
  #print(base.url)
  version.url = paste(base.url, "version", sep="/")
  cytoscape.open = TRUE
  tryCatch(expr = { GET(version.url)}, 
         error = function(e) { return (cytoscape.open = FALSE)},
         finally =function(r){ return(cytoscape.open = TRUE)})
  
  if(!cytoscape.open){
  #try and launch cytoscape
    print("Cytoscape is not open.  Please launch cytoscape.")
  } else{
    cytoscape.version =  GET(version.url)
    cy.version = fromJSON(rawToChar(cytoscape.version$content))
    print(cy.version)
  } 

  #defined threshold for GSEA enrichments (need to be strings for cyrest call)
  pvalue_gsea_threshold <- enrichmentmap_p_val
  qvalue_gsea_threshold <- enrichmentmap_q_val
  similarity_threshold <- "0.375"
  similarity_metric = "COMBINED"

  cur_model_name <- analysis_names[1]
  current_network_name <- paste(cur_model_name,pvalue_gsea_threshold,qvalue_gsea_threshold,sep="_")
  
  gsea_results_paths <- paste(gsea_output_paths,"edb",sep="/")
  gsea_results_filenames <- paste(gsea_results_paths,"results.edb",sep="/")

  # Although there is a gmt file in the gsea edb results directory it have been filtered to contain only genes represented in the expression set.  
  # If you use this fltered file you will get different pathway connectivity depending on the dataset being used.  
  # We recommend using original gmt file used for the gsea analysis and not the filtered one in the results directory.
  gmt_gsea_file <- paste(dest_gmt_file,sep="/")

  em_command = paste('enrichmentmap build analysisType="gsea" gmtFile=',gmt_gsea_file,
                   'pvalue=',pvalue_gsea_threshold, 'qvalue=',qvalue_gsea_threshold,
                   'similaritycutoff=',similarity_threshold,
                   'coefficients=',similarity_metric,
                   'ranksDataset1=',rnk_files[1],
                   'enrichmentsDataset1=',gsea_results_filenames[1], 'filterByExpressions=false',
                   'expressionDataset1=',paste(working_dir,expression_file,sep="/"),
                   sep=" ")

  #enrichment map command will return the suid of newly created network.
  response <- commandRun(em_command)

  current_network_suid <- 0
  #enrichment map command will return the suid of newly created network unless it Failed.  If it failed it will contain the word failed
  tryCatch({
    if(length(response)>0){ #fix for not getting suid
      if(grepl(pattern="Failed", response)){
        paste(response)
      } else {
        current_network_suid <- response
        response <- renameNetwork(new.name = current_network_name, network = current_network_suid, base.url)
  
        output_network_file <- paste(gsea_working_path,"screenshot_network.pdf",sep="/")
        em_command2 = paste("view export OutputFile= ",output_network_file, " options=PDF", sep="")
        commandRun(em_command2)
        em_command2 = paste("session save as file= ",paste(output_path,curr_model_name,sep="/"), sep="")
        commandRun(em_command2)
      }
    }
  }, error=function(e){output_network_file=FALSE})
}

```
`r if(gsea_section){"Click below to view GSEA results:  "}`
`r if(gsea_section){ if(output_links != ""){ output_links }}`

`r if(enrichment_map & output_network_file!=FALSE){ "Click to view EnrichmentMap screen shot: \n"}`
`r if(enrichment_map & output_network_file!=FALSE){ paste(" | [ EnrichmentMap Image](",output_subdir,"/",gsea_working_dir,"/screenshot_network.pdf)", sep="") }`  
`r if(enrichment_map & output_network_file!=FALSE){ "If multiple contrasts, load GSEA results into Cytoscape/EnrichmentMap manually."}`

