# QC and Exploratory Proteomics Analysis

`r if(!iserror && txtFolder){"## Raw Data QC Report (PTXQC)"}`
```{r PTX_QC_report, results='hide', message=FALSE, warning=FALSE}
# if txt folder provided, generate PTXQC report
tryCatch(expr={
if(txtFolder && !newcontrastonly){
  library(PTXQC)
  require(methods)
  library(data.table)

  txt_folder <- file.path(working_dir, "txt");
  report.mq <- createReport(txt_folder);
}
}, error=function(e){print("Error generating PTXQC report")} )
```
`r if(!iserror){if(txtFolder){"[ View Raw Data QC Report (PTXQC) ](txt/report_v0.92.3_MQ_pipeline.pdf)" }}` 

---

## Normalization: Boxplot and Distribution
```{r boxplots_QC_and_norm, message=FALSE, warning=FALSE, fig.width=12, fig.height=4}
tryCatch( expr={
if(!newcontrastonly){
  QCandNorm <- function(eset, norm, type='proteo', outputpath=output_plots_path,
                        annotate=annot, lookup=look_up){
    
    col_palette <- rainbow(length(levels(pData(eset)$Group)))
    annotCol <- col_palette[pData(eset)$Group]
  
    eset_matrix <- exprs(eset)
    df <- data.frame(reshape2::melt(eset_matrix, id.vars = NULL));
    colnames(df) <- c("Feature","Sample", "Intensity");
    plot1 <- ggplot(df, aes(x = Sample, y = Intensity)) + geom_boxplot(fill=annotCol) +
      theme_bw() + theme(axis.text.x=element_text(angle=45, hjust=1)) +  
      labs(title=paste(lookup[type], " log2 Intensity: Before Normalization", sep=''));
    plot2 <- ggplot(df, aes(x=Intensity, colour=Sample)) + geom_density()+ theme_bw() + 
      labs(title=paste(lookup[type], " log2 Intensity: Before Normalization", sep=''));
    
    if(norm=='quantile'){
      eset_matrix_norm <- as.matrix(normalize.quantiles(eset_matrix))
      dimnames(eset_matrix_norm) <- dimnames(eset_matrix)
    }
    else if (norm=='loess'){
      eset_matrix_norm <- as.matrix(normalizeCyclicLoess(eset_matrix, method='pairs'))
    }
    else {  eset_matrix_norm<-eset_matrix }
  
    df <- data.frame(reshape2::melt(eset_matrix_norm, id.vars = NULL));
    colnames(df) <- c("Feature","Sample", "Intensity");
    plot3 <- ggplot(df, aes(x = Sample, y = Intensity)) + geom_boxplot(fill=annotCol) +
      theme_bw() + theme(axis.text.x=element_text(angle=45, hjust=1)) + 
      labs(title=paste(lookup[type], " log2 Intensity: After ",
                       toupper(norm)," Normalization", sep=''));
    plot4 <- ggplot(df, aes(x=Intensity, colour=Sample)) + geom_density()+ theme_bw() + 
      labs(title=paste(lookup[type], " log2 Intensity: After ",
                       toupper(norm)," Normalization", sep=''));
    
    output_filename<-file.path(outputpath, paste("boxplot_histogram_",type,".pdf", sep=''));
    pdf(output_filename);
    print(plot1);
    print(plot3);
    print(plot2);
    print(plot4);
    dev.off();
  
    grid.arrange(plot1+labs(title="Raw Data"), plot3+labs(title="Normalized"),
                 plot2+theme(legend.position="none")+labs(title="Raw Data"), 
                 plot4+theme(legend.position="none")+labs(title="Normalized"),
                 top=paste(lookup[type], ":", sep=""), ncol=4);

    output_filename<-file.path(outputpath, paste("MAplots_",type,".pdf", sep=''))
    pdf(output_filename)
    for (i in 1:ncol(eset)) {
      limma::plotMA(eset_matrix, array=i, main=paste(lookup[type], "MA Plot",i,sep=" "))
      limma::plotMA(eset_matrix_norm, array=i, 
                    main=paste(lookup[type],toupper(norm),"Normalized MA Plot",i,sep=" "))
    }
    dev.off();
  
    mn <- apply(eset_matrix_norm, 1, median);
    rle <- data.frame(sweep(eset_matrix_norm, MARGIN=1, STATS=mn, FUN='-'));
    output_filename <- file.path(outputpath, paste("RLE_",type,".pdf", sep=''));
    pdf(output_filename, width=11, height=8.5);
    par(mar=c(10,4,4,2)+0.1)
    boxplot(rle, main="RLE (Relative Log Expression)\nShould be centered on 0 (blue line)",
        names=colnames(eset_matrix_norm), las=2)
    lines(x=c(0,ncol(eset_matrix_norm)+1), y=rep(0,2), col="blue", lty=2)
    lines(x=c(0,ncol(eset_matrix_norm)+1), y=rep(0.1,2), col="red", lty=2)
    par(mar=c(5,4,4,2)+0.1)
    dev.off();
     
    eset_norm <- ExpressionSet(assayData=eset_matrix_norm);
    pData(eset_norm) <- pData(eset);
    rownames(pData(eset_norm)) <- colnames(eset_matrix_norm);
    fData(eset_norm) <- fData(eset);
  
    return (eset_norm);
  }

  if(isthereProteo) {eset_global <- QCandNorm(eset_global, norm=norm_method)}
  
  if(istherePhospho) {eset_phospho <- QCandNorm(eset_phospho, norm=norm_method, type="phospho")}
}
}, error=function(e){iserror=TRUE; print("Error performing normalization");} )
```
Click below to view full size images:
`r if(!iserror){if(isthereProteo){paste(" | |  Proteomic  ](",output_plots_subdir,"/boxplot_histogram_proteo.pdf)", sep="") }}` 
`r if(!iserror){if(istherePhospho){paste(" | [  Phosphoproteomic  ](",output_plots_subdir,"/boxplot_histogram_phospho.pdf)", sep="") }}`

---

## PCA Plots
```{r PCA_plots, results='hide', fig.keep='all', fig.width=12, fig.height=4}
tryCatch(expr={
if(!newcontrastonly){
  # plot pca function (edit as needed)
  plot.pca <- function(eset, x_axis="PC1", y_axis="PC2", type="proteo", lookup=look_up, outputpath=output_plots_path) {
    
    PC_data <- prcomp(t(exprs(eset)))
    percent_variance <- summary(PC_data)$importance["Proportion of Variance",]*100
    
    pca_graph <- ggplot(data=as.data.frame(PC_data$x), aes(x=PC_data$x[,x_axis], y=PC_data$x[,y_axis], colour=pData(eset)$Group))+ 
      geom_point() + theme_bw() + theme(legend.title=element_blank()) + 
      labs(title=paste("Principal Component Analysis \nAcross All Features in All Samples - ",lookup[type],sep=""),
           x=paste(x_axis, sprintf(" (%2.0f%%)", percent_variance[x_axis]), sep=""),
           y=paste(y_axis, sprintf(" (%2.0f%%)", percent_variance[y_axis]), sep="")) +
      geom_text_repel(aes(x=PC_data$x[,x_axis], y=PC_data$x[,y_axis], label=rownames(PC_data$x)), colour="black")

    pca_loading_graph <- ggplot(data=as.data.frame(PC_data$rotation), aes(x=(PC_data$rotation[,x_axis]), y=PC_data$rotation[,y_axis])) +
      geom_text(label=rownames(PC_data$rotation), colour="black") + 
      labs(x=paste(x_axis, sep=""), y=paste(y_axis, sep=""),
           title=paste("Principal Component Factor Loadings\nAcross all Features and Samples - ",lookup[type], sep=""))
    
    output_filename <- file.path(outputpath, paste("PCAplots_",type,".pdf",sep=""))
    pdf(output_filename)
    print(pca_graph)
    print(pca_loading_graph)
    plot(PC_data, type = 'l', main=paste("Variances vs. Principal Components - ",lookup[type],sep=""))
    dev.off()
    
    return(pca_graph)
    }
  
  if(isthereProteo) { plotProteo <- plot.pca(eset_global) }

  if(istherePhospho) { plotPhospho <- plot.pca(eset_phospho, type="phospho") }

  if(isthereProteo && !istherePhospho){ grid.arrange(plotProteo, ncol=1) }
  if(!isthereProteo && istherePhospho){ grid.arrange(plotPhospho, ncol=1) }
  if(isthereProteo && istherePhospho){
    grid.arrange(plotProteo+labs(title="Proteomics"), 
                 plotPhospho+labs(title="Phospho"),ncol=2)
    }
}
}, error=function(e){iserror=TRUE; print("Error generating PCA plots");} )
```
Click below to view full size images:
`r if(!iserror){if(isthereProteo){paste(" | [  Proteomic  ](",output_plots_subdir,"/PCAplots_proteo.pdf)", sep="") }}`
`r if(!iserror){if(istherePhospho){paste(" | [  Phosphoproteomic  ](",output_plots_subdir,"/PCAplots_phospho.pdf)", sep="") }}`  

---

## Venn Diagram
```{r venn, message=FALSE, fig.width=4, fig.height=3}
tryCatch( expr={
if(isthereProteo && istherePhospho){
  hit_list <- list(Proteins=fData(eset_global)$Gene, 
                   Phosphoproteins=fData(eset_phospho)$Gene )
  fill_col <- c('green', 'blue')
  size <- rep(1,2)
  futile.logger::flog.threshold(ERROR);
  venn <- venn.diagram( x=hit_list, filename=NULL, height=2000, width=2000, 
                        fill=fill_col, cat.default.pos="outer", cat.cex=size,
                        main="Overlap of Proteins and Phosphoproteins",
                        cat.pos=c(180,180 ));
  output_filename <- file.path(output_plots_path, "VennDiagram.pdf");
  pdf(output_filename);
  grid.draw(venn);
  tmp<-dev.off();
  grid.draw(venn);
}
}, error=function(e){iserror=TRUE; print("Error generating venn diagram");} )
```

---

## Variation Plots, Filtering
```{r variation, message=FALSE, warning=FALSE}
tryCatch(expr={
if(!newcontrastonly){
  variationPlotFilter <- function(eset, type="proteo", outputpath=output_plots_path, 
                                  percent_choice=0.1) {
    emat<-exprs(eset)
    MEAN <- apply(emat, 1, mean)
    STDEV <- apply(emat, 1, sd)
    MAD <- apply(emat, 1, mad)
    MED <- apply(emat, 1, median)
    top_hits <- names(MAD)[order(MAD, decreasing=TRUE)[1:(percent_choice*length(MAD))]]
    emat_top <- emat[top_hits[top_hits!=""],]
  
    output_filename<-file.path(outputpath, paste("variation_",type,".pdf", sep=""))
    pdf(output_filename)
    plot(MEAN, STDEV, pch=".", cex=4, main=paste("mean vs. stdev: ", type, sep=""))
    plot(MED, MAD, pch=19, cex=0.5, log="", main=paste("median vs MAD: ",type, sep=""))
    points(MED[top_hits], MAD[top_hits], pch=19, cex=0.5, col="red")
    legend("topright", pch=20, col=c("black", "red"), legend=c("all points", paste("filtered top ", percent_choice*100, "%", sep="")) )
    dev.off()
    
    output_filename<-file.path(outputpath, paste("variation_",type,".jpeg", sep=""))
    jpeg(output_filename)
    pairs(emat, pch=".")
    dev.off()
    return (emat_top)
   }

  if(isthereProteo) { emat_global_top <- variationPlotFilter(eset_global) }

  if(istherePhospho) { emat_phospho_top <- variationPlotFilter(eset_phospho, type="phospho") }
  
}}, error=function(e){iserror=TRUE; print("Error generating variation plots");} )
```
Click below to view SD vs Mean plots:
`r if(!iserror){if(isthereProteo){paste(" | [  Proteomic  ](",output_plots_subdir,"/variation_proteo.pdf)", sep="") }}`
`r if(!iserror){if(istherePhospho){paste(" | [  Phosphoproteomic  ](",output_plots_subdir,"/variation_phospho.pdf)", sep="") }}`  
Click below to view Pairs plots:
`r if(!iserror){if(isthereProteo){paste(" | [  Proteomic  ](",output_plots_subdir,"/variation_proteo.jpeg)", sep="") }}`
`r if(!iserror){if(istherePhospho){paste(" | [  Phosphoproteomic  ](",output_plots_subdir,"/variation_phospho.jpeg)", sep="") }}`  

---

## Global Static Heatmaps
```{r heatmaps_qc, message=FALSE}
tryCatch(expr={
if(!newcontrastonly){
  drawthemaps <- function(eset, emat_top, type='proteo', outputpath=output_plots_path,
                        mapcolor=map_color, lookup=look_up){
  
    annotLab <- data.frame(Group = factor(pData(eset)$Group));
    annotCol <- list(Group = rainbow(length(levels(pData(eset)$Group))))
    names(annotCol$Group) <- levels(annotLab$Group)
    ha_column <- HeatmapAnnotation(df=annotLab, col=annotCol)
  
    if(mapcolor=="viridis"){mapcolor <- viridis(11); maponeway <- viridis(11);
    } else {mapcolor <- (rev(brewer.pal(11, mapcolor)));
            maponeway <- rev(brewer.pal(9, "Blues")) }
  
    output_filename <- file.path(outputpath, paste("heatmaps_",type,".pdf", sep=''));
    pdf(output_filename);
    # all features, z-score
    emat_sel <- t(scale(t(exprs(eset)))) # Z-score across rows
    emat_sel[emat_sel < -2] <- -2
    emat_sel[emat_sel > 2] <- 2
    draw(Heatmap(matrix=emat_sel, col=mapcolor, name="", top_annotation=ha_column,show_row_names=FALSE,
            column_title=paste(lookup[type], ": All features, row z score", sep='') ))
            
  
    # all features, log2 intensity
    emat_sel <- exprs(eset)
    draw(Heatmap(matrix=emat_sel, col=maponeway, name="", top_annotation=ha_column,show_row_names=FALSE,
                 column_title=paste(lookup[type], ": All proteins, log2 Intensity", sep='') ))

    # variation filter, z score
    emat_sel <- emat_top
    emat_sel <- t(scale(t(emat_sel))) # Z-score across rows
    emat_sel[emat_sel < -2] <- -2
    emat_sel[emat_sel > 2] <- 2
    draw(Heatmap(matrix=emat_sel, col=mapcolor, name="", top_annotation=ha_column,show_row_names=FALSE,
                 column_title=paste(lookup[type], ": Highest variation, row z score", sep='') ))
    tmp<-dev.off();
  }

  if(isthereProteo){ drawthemaps(eset=eset_global,emat_top=emat_global_top) }
  if(istherePhospho){ drawthemaps(eset=eset_phospho,emat_top=emat_phospho_top, type='phospho') }

}}, error=function(e){iserror=TRUE; print("Error generating static heatmaps");} )
```
Click below to view static heatmaps plots:
`r if(!iserror){if(isthereProteo){ paste(" | [  Proteomic - All  ](",output_plots_subdir,"/heatmaps_proteo.pdf)", sep="") }}`
`r if(!iserror){if(istherePhospho){ paste(" | [  Phosphoproteomic - All ](",output_plots_subdir,"/heatmaps_phospho.pdf)", sep="") }}`  

---

## Interactive Heatmap
```{r interactive_plots_qc, warning=FALSE, message=FALSE}

if(!newcontrastonly){
  makeInteractiveHM <- function(eset, type="proteo", outputpath=output_plots_path,mapcolor=map_color ){
  
  annotCol <- c("red", "green", "blue", "yellow", "green", "purple",
                "brown", "black", "grey", "orange", "white", "light green", 
                "pink", "light blue" );
  sampleCols <- annotCol[1:length(levels(pData(eset)$Group))][pData(eset)$Group];
  
  if(mapcolor=="viridis"){mapcolor <- viridis(11)
  } else {mapcolor <- colorRampPalette(rev(brewer.pal(11, mapcolor))) }

#   emat.sel <- exprs(eset[rownames(eset) %in% rownames(limmaSig),])
    output_filename <- file.path(outputpath, paste("heatmap_all_",type,".html", sep=''));
    emat_sel <- exprs(eset)
    emat_sel <- t(scale(t(emat_sel))) # Z-score across rows
    emat_sel[emat_sel < -2] <- -2
    emat_sel[emat_sel > 2] <- 2
    invisible(capture.output(tmp<-heatmaply(emat_sel, scale='none', margins=c(100,100,40,20),
              col_side_colors=pData(eset)$Group,
              colors=mapcolor,file=output_filename )));
    rm(tmp)
  }
  
  if(isthereProteo){ makeInteractiveHM(eset=eset_global) }

  if(istherePhospho){ makeInteractiveHM(eset=eset_phospho, type="phospho") }
}

```

Click below to view interactive heatmaps:
`r if(!iserror){if(isthereProteo){paste(" | [  Proteomic-All  ](",output_plots_subdir,"/heatmap_all_proteo.html)", sep="") }}`
`r if(!iserror){if(istherePhospho){paste(" | [  Phosphoroteomic-All  ](",output_plots_subdir,"/heatmap_all_phospho.html)", sep="") }}`  

---

## Save Data
```{r save_data_qc, message=FALSE }
tryCatch( expr={
# save for omics integrator, expression matrix, expression set RDS, ranked list for GSEA
saveTheFiles <- function(eset, type='proteo', outputpath=output_files_path, workingdir=working.dir,
                         saveTheEset=!newcontrastonly){
    if(saveTheEset){
      output_filename <- file.path(outputpath, paste("Data_eset_",type,".RDS", sep=""));             
      saveRDS(eset, file=output_filename); 
    }

    output_filename <- file.path(outputpath,paste("Expression_matrix_",type,".txt", sep=''));
    write.table(x=cbind(fData(eset)[,c("Gene", "Protein.names")], exprs(eset)), 
            file=output_filename, sep='\t',row.names=FALSE, col.names=TRUE, quote=FALSE);
    
    output_filename <- file.path(outputpath,paste("Uniprot_IDs",type,".txt", sep=''));
    write.table(x=fData(eset)[,"Protein"], 
            file=output_filename, sep='\t',row.names=FALSE, col.names=FALSE, quote=FALSE);
}

if(isthereProteo){ saveTheFiles(eset_global); }

if(istherePhospho){ saveTheFiles(eset_phospho, type="phospho"); }

if(newcontrastonly==FALSE && runDifferential==FALSE && saveXlsx==TRUE){   
  
  wbOut <- createWorkbook()
  
  writeDataToSheets <- function(wb, eset, mapcolor=map_color, type="proteo", lookup=look_up){
    
    annotLab <- data.frame(Group = factor(pData(eset)$Group));
    annotCol <- list(Group = gsub('.{2}$','', rainbow(length(levels(pData(eset)$Group)))))
    sampleCols <- annotCol$Group[1:length(levels(pData(eset)$Group))][pData(eset)$Group];
    mapcolor <- rev(brewer.pal(7, "RdYlBu"))
    
    stName <- unlist(lookup[type])
    addWorksheet(wb=wb, sheetName=stName)

    links <- fData(eset)$Link
    names(links) <- fData(eset)$Protein
    class(links) <- "hyperlink"

    emat_sel <- t(scale(t(exprs(eset)))) # Z-score across rows
    emat_sel[emat_sel < -2] <- -2
    emat_sel[emat_sel > 2] <- 2
    formatted_table <- data.frame(cbind(fData(eset)[,c("Gene", "Protein.names", "Protein")],emat_sel, 
                                        fData(eset)[,c("Uniprot_Function", "Uniprot_Cellular_Location", "Uniprot_Disease",
                                        "GO_biological_process", "GO_molecular_function", "GO_cellular_component", "GO_ID", "ReactomeID", "KEGG_ID")],
                                        exprs(eset) ))
    names <- make.unique(c("Gene", "Protein", "Uniprot", colnames(eset), "Uniprot_Function", "Uniprot_Cellular_Location", "Uniprot_Disease",
                           "GO_biological_process", "GO_molecular_function", "GO_cellular_component", "GO_ID", "ReactomeID", "KEGG_ID", colnames(eset) ))
    if(type=="phospho") { formatted_table <- data.frame(cbind(formatted_table, fData(eset)[,"Phospho..STY..Probabilities"]))
                          names <- c(names, "Phospho..STY..Probabilities")}
    colnames(formatted_table) <- names
    
    writeDataTable(wb=wb, sheet=stName, x=formatted_table, xy=c("A",2), keepNA=FALSE, tableStyle="TableStyleLight1")
    writeData(wb=wb, sheet=stName, x=links, xy=c("C",3))
    
    conditionalFormatting(wb=wb, sheet=stName, type="colourScale", cols=4:(3+ncol(eset)), rows=3:(2+nrow(eset)), style=mapcolor[c(1,4,7)])
    
    for (c in 1:ncol(eset)){
      addStyle(wb=wb, sheet=stName, style=openxlsx::createStyle(fgFill=sampleCols[c], textRotation=90, halign="center", valign="top"), rows=2, cols=c+3)
    }
    for (c in 1:ncol(eset)){
      addStyle(wb=wb, sheet=stName, style=openxlsx::createStyle(fgFill=sampleCols[c], textRotation=90, halign="center", valign="top"), rows=2, cols=(c+3+9+ncol(eset)) )
    }

    mergeCells(wb=wb, sheet=stName, rows=1, cols=4:(3+ncol(eset)) )
    writeData(wb=wb, sheet=stName, x="Normalized Log2 Intensity, Row Z Score", xy=c(4,1))
    addStyle(wb=wb, sheet=stName, style=openxlsx::createStyle(textDecoration="bold", halign="center"), rows=1, cols=4, stack=TRUE)
    
    mergeCells(wb=wb, sheet=stName, rows=1, cols=(13+ncol(eset)):(12+ncol(eset)+ncol(eset)) )
    writeData(wb=wb, sheet=stName, x="Normalized Log2 Intensity", xy=c((13+ncol(eset)),1))
    addStyle(wb=wb, sheet=stName, style=openxlsx::createStyle(textDecoration="bold", halign="center"), rows=1, cols=(13+ncol(eset)), stack=TRUE)

    freezePane(wb=wb, sheet=stName, firstActiveRow=3, firstActiveCol=2)
    addStyle(wb=wb, sheet=stName, style=openxlsx::createStyle(textDecoration="bold", fontColour='black'), rows=1:2, cols=1:(3+ncol(eset)+9+ncol(eset)),gridExpand=TRUE, stack=TRUE)   
    
    setRowHeights(wb=wb, sheet=stName, rows=2, heights=100)
    setColWidths(wb=wb, sheet=stName, cols=1:(3+ncol(eset)+9+ncol(eset)), widths=c(16,50,16,rep(4, ncol(eset)),50,50,50, 50,50,50, 8,8,8, rep(4, ncol(eset)) ))
  }

  if(isthereProteo){ writeDataToSheets(wb=wbOut, eset=eset_global); }

  if(istherePhospho){ writeDataToSheets(wb=wbOut, eset=eset_phospho, type="phospho"); }  
  
  output_filename=file.path(output_path, paste(project_name, "_Summary.xlsx", sep=""))
  saveWorkbook(wbOut, file=output_filename, overwrite=TRUE)
  
}

}, error=function(e){iserror=TRUE; print("Error saving data");} )
```
Click below to view expression matrix:
`r if(!iserror){if(isthereProteo){paste(" | [  Proteomic  ](",output_files_subdir,"/Expression_matrix_proteo.txt)", sep="") }}`
`r if(!iserror){if(istherePhospho){paste(" | [  Phosphoproteomic  ](",output_files_subdir,"/Expression_matrix_phospho.txt)", sep="") }}`  

---


